---
title: "Cover crop weeds GROW 2026"
author: "L Sencenbaugh"
date: "2026-06-22"
output:
  rmarkdown::html_document:
    theme: hpstr
    editor_options: 
  chunk_output_type: console
---

An analytical framework for identifying interaction effects that mediate cover crop surface residue and weed response relationships in IWM studies

Authors: Lilly Sencenbaugh 1, John M. Wallace 1, Eugene Law 2, Muthu Bagavathiannan 3, Kevin 
Bradley 4, Adam Davis 5, Wesley Everman 6, Michael Flessner 7, Nicholas Jordan 8, Lauren Lazaro 9, 
John Lindquist 10, Jason Norsworthy 11, Mark VanGessel 12, and Steven B. Mirsky 13

1 Associate Professor and Post-doctoral scholar, Plant Science Department, Pennsylvania State University, University Park, PA, USA 
2 Assistant Professor, Department of Horticulture and Crop Science, Ohio State University, Columbus, OH, USA 
3 Professor, Soil and Crop Sciences, Texas A&M University, College Station, TX 
4 Professor, Division of Plant Sciences, University of Missouri, Columbia, MO, USA 
5 Professor, Department of Crop Sciences, University of Illinois, Urbana, IL, USA 
6 Assistant Professor, Iowa State University, Department of Agronomy, Ames, IA, USA 
7 Professor, Virginia Tech, School of Plant and Environmental Sciences, Blacksburg, VA, USA 
8 Professor, Department of Agronomy and Plant Genetics, University of Minnesota, St. Paul, MN, USA
9 Assistant Professor, University of Georgia 
10 Professor, University of Nebraska 
11 Professor, University of Arkansas, Department of Crop, Soil and Environmental Sciences, Fayetteville, AK, USA
12 Professor, University of Delaware, Georgetown, DE, USA 
13 Director of Digital Agriculture, USDA-ARS Sustainable Agricultural Systems Laboratory, Beltsville, MD, USA. 


*Introduction to study*

* Managing cover crops to produce high-biomass cover crop surface residues is a promising integrated weed management strategy by reducing survival and fitness of annual weeds. 
* Here we seek to address the context dependence that influences cover crop - weed relationships that impact management recommendations. 
* We used piecewise structural equation models (pSEM) and hierarchical modeling procedures to discern the direction and magnitudes of interactive effects between cover crops, weeds, and soybean yield. 




Required packages

```{r, install packages, include = FALSE}
library(plyr) #data organization
library(dplyr) #data organization
library(ggplot2) #figures
library(lme4) #linear mixed effects models
library(lmerTest) #linear mixed effects models
library(readr) #importing csv
library(tidyr) #data organization
#install.packages("piecewiseSEM")
library(piecewiseSEM) #piecewise structural equation modeling 
library(car) #stats
#install.packages("prettydoc")
library(prettydoc) #aesthetics 
library(influence.ME) #influential points
library(broom.mixed) #random effect figures
library(effects) #effects plots
library(ggpubr) #plot stuff
library(ggeffects) #plot effects
#library(semEff) #sem effects
library(Gmisc) #plot stuff
library(scales) #plot stuff
library(grid) #plot stuff
library(MuMIn) #pairwise and group comparisons
library(lattice) # dot plot
library(merTools) #create confidence intervals from lmer models
library(performance) #checking aic and r2 values
```



#### Import the data

```{r    echo = TRUE, warning = FALSE}
#setwd("~/CCHIT 2025")
cchit <- read_csv("biomass_master.csv") 
cchit <- as.data.frame(cchit)

# setting factors and numbers
cchit$state <- as.factor(cchit$state)
cchit$year <- as.factor(cchit$year)
cchit$plot <- as.factor(cchit$plot)
cchit$term.trmt <- as.factor(cchit$term.trmt)
cchit$herb.trmt <- as.factor(cchit$herb.trmt)
cchit$region <- as.factor(cchit$region)
cchit$DPP <- as.numeric(cchit$DPP) # days pre plant (negative values, more negative means earlier termination before planting)
cchit$sy <- interaction(cchit$state, cchit$year) # site year by combining site*year
cchit$plotID <- as.factor(cchit$plotID)

#str(cchit)

cchit <- cchit %>% 
  dplyr :: rename(ccbm = "cc.bm.kgha", 
         yield = "yield.kgha", 
         weedbmh = "weed.bm.harv.kgha", 
         weedbmt = "weed.bm.term.kgha", 
         densityh = "density.harv.allspp.m2", 
         densityt = "density.term.allspp.m2", 
         sbpop = "soybean.pop.ha") #renaming 

# converting biomass in kilograms to megagrams
cchit$weedbmh_mg <- as.numeric(cchit$weedbmh/1000) 
cchit$weedbmt_mg <- as.numeric(cchit$weedbmt/1000)
cchit$ccbm_mg <- as.numeric(cchit$ccbm/1000)

# 4 weeks

itsbeen4weekssinceyoulookedatme <- read_csv("weed count data.csv") 

wap4 <- as.data.frame(itsbeen4weekssinceyoulookedatme)

# setting factors and numbers
wap4$state <- as.factor(wap4$state)
wap4$year <- as.factor(wap4$year)
wap4$block <- as.factor(wap4$block)
wap4$plot <- as.factor(wap4$plot)
wap4$term_trmt <- as.factor(wap4$term_trmt)
wap4$herb_trmt <- as.factor(wap4$herb_trmt)
wap4$wap <- as.numeric(wap4$wap)  
wap4$sy <- interaction(wap4$state, wap4$year) # site year by combining site*year
wap4$plotID <- as.factor(wap4$plotID)
wap4 <- wap4 %>%
  filter(!wap == "0") ## same thing as at termination--only 1 site year has this
str(wap4)
```

Weed density mid-season

```{r warning = FALSE}
##creating the mid season density df

sum_per <- function(x) {
  if (all(is.na(x))) NA_real_ else sum(x, na.rm = TRUE) # have to remove NAs for the log aspect AND psem gets mad if they're in there
}

quadsize <- function(x) {
  if (all(is.na(x))) NA else first(na.omit(x))
} # some site years have different plot sizes

totals <- wap4 %>% #wap = weeks after planting
  dplyr :: group_by(sy, plotID, wap, term_trmt, herb_trmt) %>%
  dplyr :: summarise(
    total_ct1 = sum_per(seq_ct_1), #frame 1
    total_ct2 = sum_per(seq_ct_2), #frame 2
    total_ct3 = sum_per(seq_qt_3 ), #frame 3
    quad1_size = quadsize(seq_quad_area_1),
    quad2_size = quadsize(seq_quad_area_2),
    quad3_size = quadsize(seq_quad_3),
    total_perm2_1 = sum_per(Count_per1m_1), #normalizing to 1m2
    total_perm2_2 = sum_per(Count_per1m_2),
    total_perm2_3 = sum_per(Count_per1m_3),
    .groups = "drop"
  )

totals <- as.data.frame(totals)

#making data frame long based on quadrat
totals <- totals %>% 
  dplyr :: rename(term.trmt = term_trmt)
totals <- totals %>%
  dplyr :: rename(herb.trmt = herb_trmt)

totals_long <- totals %>%
  pivot_longer(
    cols = c(total_perm2_1, total_perm2_2), #only texas had 3 frames
    names_to = "quadrat",
    values_to = "total_perm2"
  ) %>%
  dplyr :: mutate(
    quadrat = ifelse(quadrat == "total_perm2_1", "quad1", "quad2")
  )

totals_long <- totals_long %>%
  dplyr :: mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

totals_long <- as.data.frame(totals_long)

totals_joined <- totals_long %>% #combine!!!
  left_join(
    cchit %>% 
      dplyr::select(sy, plotID, term.trmt, herb.trmt, DPP, ccbm_mg),
    by = c("sy", "plotID", "term.trmt", "herb.trmt")
  )

totals_weedy <- totals_joined %>% filter(herb.trmt != "Standard") #weedy totals not standard

str(totals_weedy)
```


We are using the log response ratio

```{r warning = FALSE}

group_means <- totals_weedy %>%
  dplyr :: group_by(sy, plotID, term.trmt, DPP, ccbm_mg, wap, quadrat) %>%
  dplyr :: summarise(
    mean_density = mean(total_perm2, na.rm = TRUE),
    .groups = "drop"
  )

ref <- group_means %>% #reference level i.e. the denominator
  dplyr :: filter(term.trmt == "No cover") %>% 
  dplyr :: rename(ref_density = mean_density) %>% 
  dplyr :: select(sy, plotID, wap, quadrat, ref_density) #wap = weeks after plant

group_means <- group_means %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))

rr_density <- group_means %>% #combine!
  left_join(ref, by = c("sy", "block", "wap", "quadrat"), suffix = c("", "_ref")) %>% #added the _ref to tell apart for later
  mutate(
    RR = mean_density / ref_density,
    lnRR = log(RR + 0.000001) #log response ratio
  )


rr_density <- as.data.frame(rr_density)

rr_density <- rr_density %>%
  dplyr :: filter(!(wap < 2), !(wap > 10)) #MID season

str(rr_density) 
```


Density end of season

```{r warning = FALSE}
cchitnh <- subset(cchit, herb.trmt == "Weedy") # subset out herbicide treated
cchitnh <- cchitnh %>%
  dplyr :: mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

group_means <- cchitnh %>%
  dplyr :: group_by(state, year, sy, term.trmt, plotID, DPP, ccbm_mg) %>%
  dplyr :: summarize(
    mean_biomass = mean(densityh, na.rm = TRUE),
    .groups = "drop"
  ) 


ref <- group_means %>% #basically repeating above but with density at harvest!
  dplyr :: filter(term.trmt == "No cover") %>%
  dplyr :: rename(ref_biomass = mean_biomass) %>%
   dplyr :: select(sy, plotID, ref_biomass)

group_means <- group_means %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
 dplyr ::  mutate(block = substr(plotID, 2, 3))



rr_data <- group_means %>%
  left_join(ref, by = "sy", suffix = c("", "_ref")) %>%
  mutate(
    RR = mean_biomass / ref_biomass,
    lnrr1 = log(RR + 0.000001)
  )


rr_data$RR <- as.numeric(rr_data$RR)
rr_data <- as.data.frame(rr_data)
str(rr_data)

rr_data$lnrr1 <- log(rr_data$RR + 0.000001) #log!



# why are we losing site years?

#sum(is.nan(group_means$mean_biomass))
#group_means %>%
 # filter(is.nan(mean_biomass))
#ref %>%
 # filter(is.nan(ref_biomass))

# missing data (density)


```


Making a new dataset with density at harvest and 4 weeks combined 

```{r}
rr_data <- rr_data %>%
  dplyr::rename(lnRR_harvest = lnrr1) #harvest

rr_density <- rr_density %>%
 dplyr:: rename(lnRR_4WAP = lnRR) #mid

rr_data <- rr_data %>%
 dplyr:: mutate(across(c(sy, plotID, block), as.character))

rr_density <- rr_density %>%
  dplyr::mutate(across(c(sy, plotID, block), as.character))

rr_density_mean <- rr_density %>%
dplyr::  group_by(sy, block, DPP, ccbm_mg) %>%
dplyr::  summarise(lnRR_4WAP = mean(lnRR_4WAP, na.rm = TRUE), .groups = "drop")

final_data <- rr_data %>%
  dplyr :: left_join(rr_density_mean, by = c("sy", "block", "DPP", "ccbm_mg"))

str(final_data)
```


Checking for specific outliers
```{r, warning = FALSE}
ggplot(final_data, aes(x = DPP, y = lnRR_4WAP, color = state)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_smooth(method = "lm", fullrange = TRUE, size = 1, alpha = 0.2) +
  theme_light()
```

## Quantity vs Quality

For this PSEM model we have 2 component models

2. weed density 4 weeks after planting ~ cover crop biomass
1. cover crop biomass ~ days pre plant


Scaling DPP and biomass
```{r}
final_data$ccbm_sc <- scale(final_data$ccbm_mg)
final_data$DPP_sc  <- scale(final_data$DPP)
```



```{r}
mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod1a <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc || sy), data = final_data) # uncorrelated random effects
anova(mod1a, mod1) # correlated is better!
summary(mod1)

Anova(mod1, type = "III", test = "Chisq")
```


#### Table 1 

```{r}
final_data1 <- final_data %>%
  filter(!is.na(lnRR_4WAP), !is.infinite(lnRR_4WAP), !is.nan(lnRR_4WAP)) 



mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data1, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data1) # uncorrelated random effects
anova(mod2a, mod2) 


mod2c <- lm(lnRR_4WAP ~ ccbm_sc, data = final_data1 )
summary(mod2c)
AIC(mod2c)
r2(mod2c)

mod2d <- lmer(lnRR_4WAP ~ ccbm_sc + (1 | sy), data = final_data1)
summary(mod2d)
AIC(mod2d)
r2(mod2d)

mod2b <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data1) 
summary(mod2b)
r2(mod2b)
AIC(mod2b)

mod2e <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc || sy), data = final_data1) 
summary(mod2e)
r2(mod2e)
AIC(mod2e)


anova(mod2b, mod2e)

summary(mod2)
VarCorr(mod2)
Anova(mod2, type = "III", test.statistic = "Chisq")
```


## Figure 1



```{r}
state_max <- final_data %>%
  group_by(sy) %>%
  summarise(max_ccbm = max(ccbm_mg, na.rm = TRUE))

highest_common <- min(state_max$max_ccbm)

#use this for the figure 
state_year_summary <- final_data %>%
  group_by(state, year) %>%
  summarise(
    mean_ccbm = mean(ccbm_mg, na.rm = TRUE),
    median_ccbm = median(ccbm_mg, na.rm = TRUE),
    max_ccbm = max(ccbm_mg, na.rm = TRUE),
    sd_ccbm = sd(ccbm_mg, na.rm = TRUE),
    n = sum(!is.na(ccbm_mg)),
    se_ccbm = sd_ccbm / sqrt(n),
    .groups = "drop"
  )

#order from low to high by state
state_order <- state_year_summary %>%
  group_by(state) %>%
  summarise(state_mean = mean(mean_ccbm)) %>%
  arrange(state_mean) %>%
  pull(state)

state_year_summary$state <- factor(state_year_summary$state,
                                   levels = state_order)


state_year_summary$sy <- interaction(state_year_summary$state, state_year_summary$year)

```



```{r}
#how to make jitter same for the points AND error bars

jit <- position_jitter(width = 0, height = 0.2, seed = 13) #set seed to set jitter and error


yearmax <- ggplot(state_year_summary,
       aes(x = mean_ccbm, y = state, fill = state, shape = year)) +
  
  scale_shape_manual("Year", values = c(21, 22, 23, 24)) +
  scale_fill_manual("Year", values = c(   "#F3DF6C", "#CEAB07", "#FAD510","#7294D4", "#B40F20","#C52E19","#FBA72A", "#972D15","#F11B00", "#85D4E3", "#046C9A")) + 
  
  geom_errorbar(aes(xmin = mean_ccbm - se_ccbm,
                  xmax = mean_ccbm + se_ccbm),
              width = 0.3,
              position = jit) +
  geom_point(size = 1.7, position = jit) +
  labs(x = expression(Cover~crop~biomass~(Mg~ha^-1)),
       y = "State",
       fill = "state") +
  scale_x_continuous(limits = c(0, 8))+
  
  theme_light() +
  coord_fixed(ratio = 0.75) +
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))

#source("test_plot.R")
yearmax

#ggsave("yearaverages.png", width = 3.5, height = 3.5, units = "in", dpi = 300)
```


## Figure 3 a and b 

```{r}
rd1 <- broom.mixed::tidy(mod1, effects = "ran_vals", conf.int = TRUE, response = TRUE)
rd1 <- as.data.frame(rd1) 
interceptd1 <- subset(rd1, term == "(Intercept)")
sloped1 <- subset(rd1, term == "DPP_sc")
sloped1$slope <- sloped1$estimate
sloped1$slope.low <- sloped1$conf.low
sloped1$slope.high <- sloped1$conf.high
df.biplotd1 <- cbind(interceptd1, sloped1$slope, sloped1$slope.low, sloped1$slope.high)
df.biplotd1 <- as.data.frame(df.biplotd1)

wes_yellows <- c("#FAD510", "#E1BD6D", "#9A8822", "#EBCC2A", "#E1AF00", "#E3B710", "#F2AD00", "#FDD262", "#E2D200", "#F3DF6C", "#CEAB07", "#FAD77B", "#FBA72A")
wes_reds <- c("#A42820", "#9B110E", "#550307", "#CB2314", "#F2300F", "#C93312", "#EF5703", "#F11B00", "#FF0000", "#B40F20", "#972D15", "#FD6467", "#5B1A18", "#BD3027", "#C52E19")
wes_blues <- c("#273046", "#899DA4", "#35274A", "#3B9AB2", "#78B7C5", "#3A9AB2", "#6FB2C1", "#91BAB6", "#5BBCD6", "#046C9A", "#ABDDDE", "#46ACC8", "#85D4E3", "#7294D4", "#90D4CC", "#7FC0C6", "#0A9F9D", "#5785C1")


figure1b <-
  ggplot(df.biplotd1, 
         aes (x = estimate, 
              y = sloped1$slope, xmin = conf.low, 
              xmax = conf.high,  
              ymin = sloped1$slope.low, 
              ymax = sloped1$slope.high, 
              fill = level, color = level)) + 
  geom_hline(yintercept = 0, color = "gray85", size=0.75) +
  geom_vline(xintercept = 0, color = "gray85", size=0.75) +
  geom_errorbar(size=0.5, alpha = 0.3) + 
  geom_errorbarh(size=0.5, alpha = 0.3) + 
  geom_point(size = 1.5, shape = 21, alpha = 1.0, stroke = 0.8) +
  labs(x = "Conditional mode (random intercept)",
       y = "Conditional mode (random slope)", 
       color = "Site-year", fill = "Site-year") + 

  theme_light () + 
  scale_color_manual(values = c("#68180D","#68180D", #ar
                                "#266B7E", "#266B7E", "#266B7E", "#266B7E", #de
                                "#A18902", "#A18902", "#A18902", "#A18902",#il
                                "#68180D", "#68180D", "#68180D", #la
                                "#244D51", "#244D51", "#244D51", #md
                                "#AE8224", "#AE8224","#AE8224","#AE8224", #mn
                                "#AE8224", "#AE8224","#AE8224","#AE8224", #mo
                                "#68180D","#68180D","#68180D", #nc
                                "#9D8206","#9D8206", "#9D8206", "#9D8206", #ne
                                "#3F1309",  "#3F1309","#3F1309","#3F1309", #tx
                                "#266B7E", "#266B7E", "#266B7E", "#266B7E"))+ #va
  scale_fill_manual(values = c("#B40F20","#B40F20",                            #arkansas
                               "#7294D4", "#7294D4", "#7294D4", "#7294D4",    # delaware
                                "#FAD510", "#FAD510", "#FAD510",  "#FAD510",  # illinois
                                "#C52E19", "#C52E19", "#C52E19",               #louisiana
                                "#046C9A", "#046C9A", "#046C9A",              #maryland
                                "#F3DF6C", "#F3DF6C","#F3DF6C","#F3DF6C",     #minnesota
                                "#FBA72A", "#FBA72A","#FBA72A",  "#FBA72A",    #missouri
                                "#F11B00","#F11B00","#F11B00",                 #north carolina
                                "#CEAB07","#CEAB07", "#CEAB07", "#CEAB07",    #nebraska
                                "#972D15", "#972D15", "#972D15", "#972D15",   # texas
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(5, "pt"),
       legend.text = element_text(size = 7),
       legend.title = element_text(size = 8),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 6, byrow = TRUE, title.position = "top", title.hjust = 0.5))
 
figure1b

#save to make it look really pretty
#ggsave("figure1b.png", width = 6, height = 2.5, units = "in", dpi = 300)

```




```{r}
final_data2 <- final_data %>%
  filter(!is.na(DPP_sc)) 

mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data2, na.action = na.omit)
newdat1 <- expand.grid(
  DPP_sc = seq(min(final_data2$DPP_sc, na.rm = TRUE),
                max(final_data2$DPP_sc, na.rm = TRUE),
                length.out = 100),
  sy = final_data2$sy
)

newdat1$pred <- predict(mod1, newdata = newdat1,
                         re.form = NULL,
                         allow.new.levels = TRUE) # had to add this due to the missing site years in the model (that do show up in the full dataframe)

newdat1$pred_pop <- predict(mod1, newdata = newdat1, re.form = NA)
```

```{r warning = FALSE}
pred_int <- predictInterval(
  mod1,
  newdata = newdat1,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat1$fit   <- pred_int$fit
newdat1$lwr   <- pred_int$lwr
newdat1$upr   <- pred_int$upr
```

```{r}
figure1a <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat1,
            aes(x = DPP_sc, y = pred, color = sy),
            size = 0.75, alpha = 0.4) +
  
  # Population-level line (black, bold, on top)
  geom_line(data = newdat1,
            aes(x = DPP_sc, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat1,
            aes(x = DPP_sc, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
 
  scale_color_manual(values = c("#B40F20","#B40F20",                          #arkansas
                               "#7294D4", "#7294D4", "#7294D4", "#7294D4",    # delaware
                                "#FAD510", "#FAD510", "#FAD510",  "#FAD510",  # illinois
                                "#C52E19", "#C52E19", "#C52E19",               #louisiana
                                "#046C9A", "#046C9A", "#046C9A",              #maryland
                                "#F3DF6C", "#F3DF6C","#F3DF6C","#F3DF6C",     #minnesota
                                "#FBA72A", "#FBA72A","#FBA72A",  "#FBA72A",    #missouri
                                "#F11B00","#F11B00","#F11B00",                 #north carolina
                                "#CEAB07","#CEAB07", "#CEAB07", "#CEAB07",    #nebraska
                                "#972D15", "#972D15", "#972D15", "#972D15",   # texas
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
  
  
  labs(
    y = "CCSR biomass", 
    x = "Termination timing", 
       color = "Site-year") + 
  theme_light()+
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none")


figure1a

#save to make it look really pretty
#ggsave("figure1a.png", width = 6, height = 6, units = "in", dpi = 300)

figure1 <- ggarrange(figure1a, figure1b)
figure1


#ggsave("figure1.png", width = 6, height = 2.5, units = "in", dpi = 300)
```


Resetting everything without the site-years missing density data

Weed density mid-season

```{r warning = FALSE}
group_means <- totals_weedy %>%
  dplyr :: group_by(sy, plotID, term.trmt, DPP, ccbm_mg, wap, quadrat) %>%
  dplyr :: summarise(
    mean_density = mean(total_perm2, na.rm = TRUE),
    .groups = "drop"
  )

ref <- group_means %>% 
  dplyr :: filter(term.trmt == "No cover") %>% 
  dplyr :: rename(ref_density = mean_density) %>% 
  dplyr :: select(sy, plotID, wap, quadrat, ref_density)

group_means <- group_means %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))

rr_density <- group_means %>%
  dplyr :: left_join(ref, by = c("sy", "block", "wap", "quadrat"), suffix = c("", "_ref")) %>%
  dplyr :: filter(!is.na(ref_density), !is.na(mean_density)) %>%
  dplyr :: mutate(
    RR = ifelse(ref_density == 0, NA, mean_density / ref_density),
    lnRR = log(RR + 0.000001) #psem needs help interpretting log (can't do it within psem code must be done prior)
  ) %>%
  dplyr::select(-plotID_ref) %>% # keep the main columns tidy 
  filter(!is.na(lnRR), !is.infinite(lnRR), !is.nan(lnRR)) 

rr_density <- as.data.frame(rr_density)



str(rr_density)
```
Density at harvest

```{r warning = FALSE}
cchitnh <- subset(cchit, herb.trmt == "Weedy") # subset out herbicide treated
cchitnh <- cchitnh %>%
  dplyr::mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

group_means <- cchitnh %>%
  dplyr::group_by(state, year, sy, term.trmt, plotID, DPP, ccbm_mg) %>%
  dplyr::summarize(
    mean_biomass = mean(densityh, na.rm = TRUE),
    .groups = "drop"
  ) 

ref <- group_means %>%
  dplyr::filter(term.trmt == "No cover") %>%
  dplyr::rename(ref_biomass = mean_biomass) %>%
   dplyr :: select(sy, plotID, ref_biomass)

group_means <- group_means %>%
  dplyr::mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
  dplyr::mutate(block = substr(plotID, 2, 3))

rr_data <- group_means %>%
  dplyr::left_join(ref, by = "sy", suffix = c("", "_ref")) %>%
  dplyr::filter(!is.na(ref_biomass), !is.na(mean_biomass)) %>%
  dplyr::mutate(RR = mean_biomass / ref_biomass) %>%
  dplyr::select(-plotID_ref, -block_ref) 

rr_data[sapply(rr_data, is.infinite)] <- NA
rr_data <- na.omit(rr_data)


rr_data$RR <- as.numeric(rr_data$RR)
rr_data <- as.data.frame(rr_data)
str(rr_data)

rr_data$lnrr <- log(rr_data$RR + 0.000001) 
```

Making a new dataset with density at harvest and 4 weeks combined 

```{r}
rr_data <- rr_data %>%
  dplyr::rename(lnRR_harvest = lnrr)

rr_density <- rr_density %>%
 dplyr:: rename(lnRR_4WAP = lnRR)

rr_data <- rr_data %>%
 dplyr:: mutate(across(c(sy, plotID, block), as.character))

rr_density <- rr_density %>%
  dplyr::mutate(across(c(sy, plotID, block), as.character))

rr_density_mean <- rr_density %>%
dplyr::  group_by(sy, block, DPP, ccbm_mg) %>%
dplyr::  summarise(lnRR_4WAP = mean(lnRR_4WAP, na.rm = TRUE), .groups = "drop")

final_data <- rr_data %>%
  dplyr :: left_join(rr_density_mean, by = c("sy", "block", "DPP", "ccbm_mg"))

final_data$sy <- as.factor(final_data$sy)
str(final_data)
```


Removing specific outliers
```{r, warning = FALSE}
#x11()
#plot(final_data$DPP, final_data$lnRR_4WAP, pch=19, cex=2, col="goldenrod")

# identify points
#idx <- identify(final_data$DPP, final_data$lnRR_4WAP, labels = final_data$plotID)

# confirm
#final_data[idx, c("plotID", "DPP", "lnRR_harvest")]

# remove by row

final_data <- subset(final_data, !(final_data$plotID %in% c("M73MW", "D73LW", "V02LW")))

ggplot(final_data, aes(x = DPP, y = lnRR_4WAP, color = state)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_smooth(method = "lm", fullrange = TRUE, size = 1, alpha = 0.2) +
  theme_light()
```

Scaling

```{r}
final_data$ccbm_sc <- scale(final_data$ccbm_mg)
final_data$DPP_sc  <- scale(final_data$DPP)
```

### Model 1 (cover crop biomass ~ days pre planting)

```{r}
mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod1a <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc || sy), data = final_data) # uncorrelated random effects
anova(mod1a, mod1) # correlated is better!
summary(mod1)
Anova(mod1, type = "III", test = "Chisq")
```


```{r   fig.width = 5, fig.height = 5}
residuals <- resid(mod1)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod1, type="link"), pearson = residuals(mod1, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod1, vars = "Cook")
dotplot(ranef(mod1))
```

```{r, fig.height = 15, fig.width = 10}
re <- ranef(mod1)$sy
re_df <- re %>%
  tibble::rownames_to_column("sy") %>%
  rename(rand_intercept = `(Intercept)`)
fix_int <- fixef(mod1)[1]
re_df <- re_df %>%
  mutate(intercept = fix_int + rand_intercept)
final_data1 <- final_data %>%
  left_join(re_df, by = "sy")
final_data1 <- final_data1 %>%
  mutate(sy = reorder(sy, intercept))

ggplot(final_data1, aes(x = DPP_sc.x, y = ccbm_sc)) +
  geom_point(size = 2, alpha = 0.2, pch = 16) +
  geom_smooth(method = "lm", se = TRUE,
              fullrange = TRUE, 
              size = 0.75, alpha = 0.5,
              col = "hotpink") +
  theme_light() + 
  theme(axis.title = element_text(size = 14),
        legend.title = element_text(size = 14), 
        aspect.ratio = 1.5)+
  labs(y = "cover crop biomass scaled", x = "days pre planting")+
  facet_wrap(~sy)
```



### Model 2 (lnRR ~ ccbm)


```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated

summary(mod2)
VarCorr(mod2)
Anova(mod2, type = "III", test.statistic = "Chisq")
```


```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data) # uncorrelated random effects
anova(mod2a, mod2) # correlated is better!

summary(mod2)
VarCorr(mod2)
Anova(mod2, type = "III", test.statistic = "Chisq")
```

```{r   fig.width = 5, fig.height = 5}
residuals <- resid(mod2)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod2, type="link"), pearson = residuals(mod2, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod2, vars = "Cook")
dotplot(ranef(mod2))
```

How do weeds decline as cover crop biomass goes up?

```{r}
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))

mod <- lm(lnRR_4WAP ~ ccbm_sc , data= final_data, na.action = na.omit)
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc || sy), data = final_data) # uncorrelated random effects

#previous version included DPP, removed here

#first make 2Mg in the scaled scale
ccbm_mean <- mean(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sd   <- sd(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sc_2mg <- (2 - ccbm_mean) / ccbm_sd #finding 2 but scaled


newdat <- data.frame(
  ccbm_sc = ccbm_sc_2mg
)

pred_lnRR <- predict(mod, newdata = newdat) #predict
RR <- exp(pred_lnRR) #back transform the lnrr 
percent_reduction <- (1 - RR) * 100 #percent reduction

# mixed model just intercept

mod2c <- lmer(lnRR_4WAP ~ ccbm_sc  + ( 1 | sy), data = final_data) # uncorrelated random effects

pred_lnRR_int <- predict(mod2c, newdata = newdat, re.form = NA)
RR_int <- exp(pred_lnRR_int) #back transform the lnrr 
percent_reduction_int <- (1 - RR_int) * 100 #percent reduction



#correlated mixed model
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc +  (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated


pred_lnRR_corr <- predict(mod2, newdata = newdat, re.form = NA)
RR_corr <- exp(pred_lnRR_corr) #back transform the lnrr 
percent_reduction_corr <- (1 - RR_corr) * 100 #percent reduction

```



1Mg

```{r}
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))

mod <- lm(lnRR_4WAP ~ ccbm_sc , data= final_data, na.action = na.omit)
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc || sy), data = final_data) # uncorrelated random effects



#first make 1Mg in the scaled scale
ccbm_mean <- mean(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sd   <- sd(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sc_1mg <- (1 - ccbm_mean) / ccbm_sd #finding 2 but scaled

newdat <- data.frame(
  ccbm_sc = ccbm_sc_1mg
 
)

pred_lnRR <- predict(mod, newdata = newdat) #predict
RR <- exp(pred_lnRR) #back transform the lnrr 
percent_reduction <- (1 - RR) * 100 #percent reduction

# mixed model just intercept

mod2c <- lmer(lnRR_4WAP ~ ccbm_sc  + ( 1 | sy), data = final_data) # uncorrelated random effects

pred_lnRR_int <- predict(mod2c, newdata = newdat, re.form = NA)
RR_int <- exp(pred_lnRR_int) #back transform the lnrr 
percent_reduction_int <- (1 - RR_int) * 100 #percent reduction



#correlated mixed model
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))
final_data$DPP_sc  <- as.numeric(scale(final_data$DPP))
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc +  (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated


pred_lnRR_corr <- predict(mod2, newdata = newdat, re.form = NA)
RR_corr <- exp(pred_lnRR_corr) #back transform the lnrr 
percent_reduction_corr <- (1 - RR_corr) * 100 #percent reduction

```



4Mg

```{r}
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))

mod <- lm(lnRR_4WAP ~ ccbm_sc , data= final_data, na.action = na.omit)
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc || sy), data = final_data) # uncorrelated random effects



#first make 4Mg in the scaled scale
ccbm_mean <- mean(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sd   <- sd(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sc_4mg <- (4 - ccbm_mean) / ccbm_sd #finding 2 but scaled


newdat <- data.frame(
  ccbm_sc = ccbm_sc_4mg
)

pred_lnRR <- predict(mod, newdata = newdat) #predict
RR <- exp(pred_lnRR) #back transform the lnrr 
percent_reduction <- (1 - RR) * 100 #percent reduction

# mixed model just intercept

mod2c <- lmer(lnRR_4WAP ~ ccbm_sc  + ( 1 | sy), data = final_data) # uncorrelated random effects

pred_lnRR_int <- predict(mod2c, newdata = newdat, re.form = NA)
RR_int <- exp(pred_lnRR_int) #back transform the lnrr 
percent_reduction_int <- (1 - RR_int) * 100 #percent reduction



#correlated mixed model
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))
final_data$DPP_sc  <- as.numeric(scale(final_data$DPP))
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc +  (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated


pred_lnRR_corr <- predict(mod2, newdata = newdat, re.form = NA)
RR_corr <- exp(pred_lnRR_corr) #back transform the lnrr 
percent_reduction_corr <- (1 - RR_corr) * 100 #percent reduction
```



### PSEM

Starting-- most basic model

DPP -> ccbm 
ccbm -> lnrr_mid
```{r}
psemDensity <- psem(
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity, .progressBar = FALSE)
# p = 0.059
fisherC(psemDensity)
#Fisher's: 5.643 on 2 df, p = 0.06	
summary(psemDensity)
```

The p = 0.059 is very borderline. Moderate evidence of a missing pathway. 

DPP -> ccbm 
ccbm -> lnrr_mid
```{r}
psemDensity1 <- psem(
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity1, .progressBar = FALSE)
# saturated
fisherC(psemDensity1)
#Fisher's: 5.643 on 2 df, p = 0.06	
summary(psemDensity1)
```


Saturated model



## Figure 3 c and d


```{r}
mod2b <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data) # remove dpp
summary(mod2b)
r2(mod2b)
AIC(mod2b)
```

```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data) # remove dpp
rd2 <- broom.mixed::tidy(mod2, effects = "ran_vals", conf.int = TRUE, response = TRUE)
rd2 <- as.data.frame(rd2) 
interceptd2 <- subset(rd2, term == "(Intercept)")
sloped2 <- subset(rd2, term == "ccbm_sc")
sloped2$slope <- sloped2$estimate
sloped2$slope.low <- sloped2$conf.low
sloped2$slope.high <- sloped2$conf.high
df.biplotd2 <- cbind(interceptd2, sloped2$slope, sloped2$slope.low, sloped2$slope.high)
df.biplotd2 <- as.data.frame(df.biplotd2)

figure1d <-
  ggplot(df.biplotd2, 
         aes (x = estimate, 
              y = sloped2$slope, xmin = conf.low, 
              xmax = conf.high,  
              ymin = sloped2$slope.low, 
              ymax = sloped2$slope.high, 
              fill = level, color = level)) + 
  geom_hline(yintercept = 0, color = "gray85", size=0.75) +
  geom_vline(xintercept = 0, color = "gray85", size=0.75) +
  geom_errorbar(size=0.5, alpha = 0.3) + 
  geom_errorbarh(size=0.5, alpha = 0.3) + 
  geom_point(size = 1.5, shape = 21, alpha = 1.0, stroke = 0.8) +
  labs(x = "Conditional mode (random intercept)",
       y = "Conditional mode (random slope)", 
       color = "Site-year", fill = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#266B7E", "#266B7E",  
                                "#A18902", "#A18902", 
                                "#68180D", "#68180D", 
                                "#023146", 
                                "#AE8224",
                                "#AE8224", "#AE8224",
                                "#9D8206", "#9D8206","#9D8206",  
                                "#266B7E", "#266B7E", "#266B7E", "#266B7E"))+

   scale_fill_manual(values = c("#7294D4", "#7294D4",  #delaware
                                "#FAD510", "#FAD510", #illinois
                                "#C52E19", "#C52E19", #louisiana
                                "#046C9A", #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A",#missouri
                                "#CEAB07", "#CEAB07", "#CEAB07", #nebraska
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) +
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))




figure1d
#ggsave("figure1d.png", width = 7, height = 2.5, units = "in", dpi = 300)
```



```{r}

mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data) # remove dpp

# am having issues with dropped site years sneaking back into df
# in the dataframe we have lots of site years but in mod2 only 17 have enough observations

valid_sy <- unique(mod2@frame$sy) #just site years in mod2

newdat <- expand.grid(
  ccbm_sc = seq(min(final_data$ccbm_sc, na.rm = TRUE),
                max(final_data$ccbm_sc, na.rm = TRUE),
                length.out = 100),
  DPP_sc = mean(final_data$DPP_sc, na.rm = TRUE),
  sy = valid_sy
)
newdat$pred_sy <- predict(mod2, newdata = newdat, re.form = NULL)

newdat$pred_pop <- predict(mod2, newdata = newdat, re.form = NA)
```


Confidence interval

```{r}
pred_int1 <- predictInterval(
  mod2,
  newdata = newdat,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat$fit   <- pred_int1$fit
newdat$lwr   <- pred_int1$lwr
newdat$upr   <- pred_int1$upr
```


```{r}
figure1c <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_sy, color = sy),
            size = 0.75, alpha = 0.4) +
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat,
            aes(x = ccbm_sc, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
  labs(x = "CCSR biomass",
       y = "Weed density (LRR; 4 WAP)", 
       color = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#7294D4", "#7294D4", #delaware
                                "#FAD510", "#FAD510", #illinois
                                "#C52E19", "#C52E19", #louisiana
                                "#046C9A", #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A",#missouri
                                "#CEAB07", "#CEAB07", "#CEAB07", #nebraska
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
 
   
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
        legend.position = "none")





figure1bottom <- ggarrange(figure1c, figure1d)
figure1bottom

#work out some save details 
#ggsave("figure1bottom.png", width = 6, height = 2.5, units = "in", dpi = 300)
```



### Figure 4b

```{r}
#the scaling
#final_data$ccbm_sc <- scale(final_data$ccbm_mg)
#final_data$DPP_sc  <- scale(final_data$DPP)
#the model
#mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)

ccbm_mean <- mean(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sd   <- sd(final_data$ccbm_mg, na.rm = TRUE)

DPP_mean <- mean(final_data$DPP, na.rm = TRUE)
DPP_sd   <- sd(final_data$DPP, na.rm = TRUE)

ccbm_vals <- c(2) #common highest mean among sites yikes
ccbm_sc_vals <- (ccbm_vals - ccbm_mean) / ccbm_sd #un-scale 

DPP_val <- -14 #focus on 2 weeks pre planting
DPP_sc_val <- (DPP_val - DPP_mean) / DPP_sd #un-scale

model_sy <- unique(mod2@frame$sy) #site year

newdata <- expand.grid(
  sy = model_sy,
  ccbm_sc = ccbm_sc_vals,
  DPP_sc = DPP_sc_val
)

newdata$biomass <- rep(c("2 Mg/ha"), 
                       each = length(model_sy))

```



```{r}
newdata$fit_sy <- predict(mod2, newdata = newdata, re.form = NULL)
newdata$fit_pop <- predict(mod2, newdata = newdata, re.form = NA)
pred_pop <- predictInterval(
  mod2,
  newdata = newdata,
  level = 0.95,
  n.sims = 2000,
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"
)
# convert to %
newdata$percent_sy  <- (1 - exp(newdata$fit_sy)) * 100
newdata$percent_pop <- (1 - exp(newdata$fit_pop)) * 100

newdata$pop_lwr <- (1 - exp(pred_pop$upr)) * 100
newdata$pop_upr <- (1 - exp(pred_pop$lwr)) * 100

pop_data <- data.frame(
  sy = "Population",
  percent_pop = unique(newdata$percent_pop)[1],
  pop_lwr = unique(newdata$pop_lwr)[1],
  pop_upr = unique(newdata$pop_upr)[1]
)

pop_mean <- unique(newdata$percent_pop)[1]

plot_data <- newdata %>%
  mutate(
    deviation = percent_sy - pop_mean
  )

plot_data <- plot_data %>%
  arrange(percent_sy)

plot_data$sy <- factor(plot_data$sy, levels = plot_data$sy)
```


```{r}
figure2b1 <- ggplot(plot_data, aes(y = sy, fill = sy, color = sy)) +
 geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "gray50",
             size = 0.25) + 
  # segment from population mean to site-year
  geom_segment(aes(x = pop_mean, xend = percent_sy,
                   yend = sy),
               color = "gray20", size = 0.5) +
  
  
  # population reference line
  geom_vline(xintercept = pop_mean,
             linetype = "longdash",
             color = "black", 
             size = 0.5) +
  # population reference point
 geom_point(data = pop_data,
           aes(x = percent_pop, y = sy),
           shape = 21,
           size = 3,
           fill = "black",
           color = "black") + 
  # population reference point error bar
geom_errorbarh(data = pop_data,
               aes(xmin = pop_lwr,
                   xmax = pop_upr,
                   y = sy),
               height = 0.5,
               color = "black",
               size = 0.5) +

    # site-year points
  geom_point(aes(x = percent_sy),
             shape = 21, size = 1.25, stroke = 0.7) +
  
    
  scale_color_manual(values = c("#9D8206", #NE
                                "#AE8224",#MN
                                "#AE8224",#MO
                                "#A18902",#IL
                                "#266B7E",#VA
                                "#023146", #DE
                                "#A18902", #IL
                                "#266B7E", #VA
                                "#266B7E",#VA
                                "#023146",#MD
                                "#9D8206",#NE
                                "#9D8206",#NE
                                "#AE8224",#MO
                                "#266B7E",#VA
                                "#68180D", #LA
                                "#023146",#DE
                                "#68180D", #LA
                                "black"))+
  scale_fill_manual(values = c( "#CEAB07", #NE
                               "#F3DF6C", #MN
                               "#FBA72A", #MO
                               "#FAD510", #IL
                               "#85D4E3",  #VA
                               "#7294D4", #DE
                               "#FAD510", #IL
                               "#85D4E3", #VA
                               "#85D4E3", #VA
                               "#046C9A", #MD
                               "#CEAB07", 
                               "#CEAB07", #NE
                               "#FBA72A", #MO
                               "#85D4E3", #VA
                               "#C52E19", #LA
                               "#7294D4",#DE
                               "#C52E19", #LA
                               "black" #population
                               ))+
  
  labs(x = "% reduction in weed density",
       y = "Site-Year") +
  
  theme_light() +
  theme(
       axis.text.x = element_text(size=8), 
        axis.text.y = element_text(size=7.5, margin = margin(r = -0.5)),
        axis.title.y = element_text(size=10, margin = margin(r = -0.5)), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))
figure2b1
ggsave("percentdiff2mgall.png", width = 2.5, height = 2.5, units = "in", dpi = 300)
```


### Figure 4c
```{r}
dat <- na.omit(final_data[, c("lnRR_4WAP", "ccbm_sc", "DPP_sc")])

y_resid <- resid(lm(lnRR_4WAP ~ ccbm_sc, data = dat))
x_resid <- resid(lm(DPP_sc ~ ccbm_sc, data = dat))

plot(x_resid, y_resid)

plot <- ggplot(dat, aes(x = x_resid, y = y_resid)) +
  geom_point(data = dat, aes(x = x_resid, y = y_resid),
            shape = 16, col = "gray", size = 2, alpha = 0.25)+


#geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.15) + 
  geom_hline(yintercept = 0, lty = "dashed", size = 0.5)+
  geom_smooth(method = "lm", se = TRUE, 
          size = 1.2, col = "gray45")+
labs(
  x = "Cover crop termination timing",
  y = "Weed density (LRR; 4 WAP)") + 
 theme_light() +
  theme(
       axis.text.x = element_text(size=8), 
        axis.text.y = element_text(size=7.5, margin = margin(r = -0.5)),
        axis.title.y = element_text(size=10, margin = margin(r = -0.5)), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))
plot
ggsave("residuals.png", width = 2.5, height = 2.5, units = "in", dpi = 300)
```


## Temporal dynamics

I will be repeating the above process for the next pSEM which includes mid- and end- of season weed density

There will be 3 component models

3. cover crop biomass ~ termination timing
2. Weed density mid season ~ cover crop biomass
1. Weed density end season ~ Weed density mid season

Removing specific outliers
```{r}
#x11()
#plot(final_data$DPP, final_data$lnRR_harvest, pch=19, cex=2, col="goldenrod")

# identify points
#idx <- identify(final_data$DPP, final_data$lnRR_harvest, labels = final_data$plotID)

# confirm
#final_data[idx, c("plotID", "DPP", "lnRR_harvest")]

# remove by row

final_data <- subset(final_data, !(final_data$plotID %in% c("L92LW", "D93LW", "M86LW", "M04LW")))

ggplot(final_data, aes(x = DPP, y = lnRR_4WAP, color = state)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_smooth(method = "lm", fullrange = TRUE, size = 1, alpha = 0.2) +
  theme_light()
```


### Model 3 (cover crop biomass ~ days pre planting)

Same as above model 1

```{r eval = FALSE}
mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod1a <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc || sy), data = final_data) # uncorrelated random effects
anova(mod1a, mod1) # correlated is better!
summary(mod1)
```

### Model 4 (weeds mid ~ cover crop biomass)

Same as above model 2


```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data) # uncorrelated random effects
anova(mod2a, mod2) # correlated is better!

mod2b <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data) # remove dpp
anova(mod2b, mod2) # on the edge here

summary(mod2)
```

### Model 5 (weeds end ~ weeds mid)


```{r}
mod5 <- lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)
#slope correlated
#singularity
anova(mod5) # non significant ccbm fixed
mod5b <- lmer(lnRR_harvest ~ lnRR_4WAP + DPP_sc + (ccbm_sc | sy),data = final_data) # remove ccbm
anova(mod5b, mod5) # no difference
anova(mod5b)

mod5 <- lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy),data = final_data) 
mod5a <- lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data) # uncorrelated random effects
#no singularity
anova(mod5a, mod5) # correlated is better!
#except we're still having singularities (but not lack of convergence)

summary(mod5)
```


```{r   fig.width = 5, fig.height = 5}
residuals <- resid(mod5)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod5, type="link"), pearson = residuals(mod5, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod5, vars = "Cook")
dotplot(ranef(mod5))
```



### PSEM


DPP -> ccbm 
ccbm -> lnrr_mid
lnrr_mid -> lnrr_end
```{r}
psemDensity2 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity2, .progressBar = FALSE)
#  p values v small
fisherC(psemDensity2)
#Fisher's: 15.98 on 6 df, p = 0.014
summary(psemDensity2)
```

```{r}
psemDensity4 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity4, .progressBar = FALSE)
# still small p value
fisherC(psemDensity4)
# Fisher's C 12.729	on 4 df p = 0.013	
summary(psemDensity4)
```



```{r}
psemDensity8 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity8, .progressBar = FALSE)
# 0.05
fisherC(psemDensity8)
# Fisher's C 5.924	on 2 df p = 0.05	
summary(psemDensity8)
```

Here we have another borderline dsep test. P = 0.05

Fully saturated model check

```{r}
psemDensity7 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc +(ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity7, .progressBar = FALSE)
# saturated
fisherC(psemDensity7)
# saturated
summary(psemDensity7)
```



## Cash crop presence


Soybean yield

Starting everything from scratch

```{r warning = FALSE, include = FALSE}
#setwd("~/CCHIT 2025")
cchit <- read_csv("biomass_master.csv") 
cchit <- as.data.frame(cchit)

cchit$state <- as.factor(cchit$state)
cchit$year <- as.factor(cchit$year)
cchit$plot <- as.factor(cchit$plot)
cchit$term.trmt <- as.factor(cchit$term.trmt)
cchit$herb.trmt <- as.factor(cchit$herb.trmt)
cchit$region <- as.factor(cchit$region)
cchit$DPP <- as.numeric(cchit$DPP) 
cchit$sy <- interaction(cchit$state, cchit$year) 
cchit$plotID <- as.factor(cchit$plotID)

str(cchit)

cchit <- cchit %>% dplyr::rename(ccbm = "cc.bm.kgha", yield = "yield.kgha", weedbmh = "weed.bm.harv.kgha", weedbmt = "weed.bm.term.kgha", densityh = "density.harv.allspp.m2", densityt = "density.term.allspp.m2", sbpop = "soybean.pop.ha") 

cchit$weedbmh_mg <- as.numeric(cchit$weedbmh/1000) 
cchit$weedbmt_mg <- as.numeric(cchit$weedbmt/1000)
cchit$ccbm_mg <- as.numeric(cchit$ccbm/1000)

cchith <- subset(cchit, herb.trmt == "Weedy") # subset 
cchith$yield_mg <- as.numeric(cchith$yield/1000)
cchith <- cchith %>%
 dplyr:: mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

group_means <- cchith %>%
dplyr::  group_by(state, year, sy, term.trmt, plotID, DPP, ccbm_mg) %>%
  dplyr::summarise(mean_yield = mean(yield_mg, na.rm = TRUE), .groups = "drop")

group_means <- group_means %>%
  dplyr::mutate(block = substr(plotID, 2, 3))

ref <- group_means %>%
  dplyr::filter(term.trmt == "No cover") %>%
  dplyr::rename(ref_biomass = mean_yield) %>%
  dplyr::select(sy, block, ref_biomass)

rr_yield <- group_means %>%
  left_join(ref, by = c("sy", "block")) %>%
  mutate(
    RR = mean_yield / ref_biomass,
    lnyield = log(mean_yield + 1e-6)
  ) %>%
  filter(!is.na(RR), is.finite(RR))

  
rr_yield[sapply(rr_yield, is.infinite)] <- NA
rr_yield <- na.omit(rr_yield)


rr_yield$RR <- as.numeric(rr_yield$RR)
rr_yield <- as.data.frame(rr_yield)

rr_yield$lnyield <- log(rr_yield$mean_yield + 0.000001) 

str(rr_yield)

```


Density at harvest

```{r warning = FALSE}
cchitnh <- subset(cchit, herb.trmt == "Weedy") # subset out herbicide treated
cchitnh <- cchitnh %>%
  dplyr::mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

group_means <- cchitnh %>%
  dplyr::group_by(state, year, sy, term.trmt, plotID, DPP, ccbm_mg) %>%
  dplyr::summarize(
    mean_biomass = mean(densityh, na.rm = TRUE),
    .groups = "drop"
  ) 

ref <- group_means %>%
  dplyr::filter(term.trmt == "No cover") %>%
  dplyr::rename(ref_biomass = mean_biomass) %>%
   dplyr :: select(sy, plotID, ref_biomass)

group_means <- group_means %>%
  dplyr::mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
  dplyr::mutate(block = substr(plotID, 2, 3))

rr_data <- group_means %>%
  dplyr::left_join(ref, by = "sy", suffix = c("", "_ref")) %>%
  dplyr::filter(!is.na(ref_biomass), !is.na(mean_biomass)) %>%
  dplyr::mutate(RR = mean_biomass / ref_biomass) %>%
  dplyr::select(-plotID_ref, -block_ref) 

rr_data[sapply(rr_data, is.infinite)] <- NA
rr_data <- na.omit(rr_data)


rr_data$RR <- as.numeric(rr_data$RR)
rr_data <- as.data.frame(rr_data)
str(rr_data)

rr_data$lnrr <- log(rr_data$RR + 0.000001) 
```



Making a new dataset with all combined

```{r}
#rr_yield <- rr_yield %>%
 #rename(lnyield = lnrr)

#rr_density <- rr_density %>%
 # rename(lnRR_4WAP = lnRR)

rr_data <- rr_data %>%
  dplyr::rename(lnRR_harvest = lnrr)

rr_yield <- rr_yield %>%
  dplyr::mutate(across(c(sy, plotID, block), as.character))

#rr_density <- rr_density %>%
 # mutate(across(c(sy, plotID, block), as.character))

rr_data <- rr_data %>%
  dplyr::mutate(across(c(sy, plotID, block), as.character))

rr_density_mean <- rr_density %>%
  dplyr::group_by(sy, block, DPP, ccbm_mg) %>%
  dplyr::summarise(lnRR_4WAP = mean(lnRR_4WAP, na.rm = TRUE), .groups = "drop")

rr_yield_mean <- rr_yield %>%
  dplyr::group_by(sy, block, DPP, ccbm_mg) %>%
  dplyr::summarise(lnyield = mean(lnyield, na.rm = TRUE), .groups = "drop")

final_datay <- rr_data %>%
  left_join(rr_density_mean, by = c("sy", "block", "DPP", "ccbm_mg"))
final_datay <- final_datay %>%
  left_join(rr_yield_mean, by = c("sy", "block", "DPP", "ccbm_mg"))

str(final_datay)
```


### Model 9
```{r}
final_datay$ccbm_sc <- as.numeric(scale(final_datay$ccbm_mg))
final_datay$DPP_sc  <- as.numeric(scale(final_datay$DPP))
```

```{r}
mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
```

## Model 10
```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc | sy), data = final_datay, na.action = na.omit)
```

## Model 11 (lnrr harvest ~ lnrr mid)

```{r}
mod5a <- lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)
```

## Model 11.5 (lnrr harvest ~ ccbm)


```{r}
mod11_5 <- lmer(lnRR_harvest ~ ccbm_sc  + (ccbm_sc || sy), data = final_datay)
mod11_5a <- lmer(lnRR_harvest ~ ccbm_sc  + (ccbm_sc | sy), data = final_datay)
anova(mod11_5, mod11_5a) # correlated is way better!
summary(mod11_5a)
anova(mod11_5a)
```



## Model 12.5 (soybean yield ~ lnrr end)

```{r}
mod9a <- lmer(lnyield ~ lnRR_harvest + lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
anova(mod9a) #biggest p value is 4wap
mod9b <- lmer(lnyield ~ lnRR_harvest +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
mod9c <- lmer(lnyield ~ lnRR_harvest +  ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)

anova(mod9b, mod9c)
```

P-value >0.05 as such, best model is the most parsimonious.


```{r   fig.width = 5, fig.height = 5}
residuals <- resid(mod9c)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod9c, type="link"), pearson = residuals(mod9c, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod9c, vars = "Cook")
dotplot(ranef(mod9c))
```





## Model 12 (soybean yield ~ lnrr 4wap)

```{r}
mod9a <- lmer(lnyield ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
anova(mod9a) #biggest p value is 4wap
mod9b <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
mod9c <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)

anova(mod9b, mod9c)
```

P-value >0.05 as such, best model is the most parsimonious.


### PSEM

Only using weed density end-of-season


```{r}
psemDensity2 <- psem(
  lmer(lnyield ~ lnRR_harvest + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_harvest ~ ccbm_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity2, .progressBar = FALSE)
# 2 p values v small
fisherC(psemDensity2)
#Fisher's: 55.917	on 6 df, p = 0
summary(psemDensity2)
```



```{r}
psemDensity3 <- psem(
  lmer(lnyield ~ lnRR_harvest + ccbm_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_harvest ~ ccbm_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity3, .progressBar = FALSE)
# still small p values
fisherC(psemDensity3)
#fishers c = 52.409, p < 0.001
summary(psemDensity3)
```


```{r}
psemDensity5 <- psem(
  lmer(lnyield ~ lnRR_harvest + ccbm_sc +  DPP_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_harvest ~ ccbm_sc +  DPP_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity5, .progressBar = FALSE)
# saturated
fisherC(psemDensity5)
#saturated
summary(psemDensity5)
```


### Figure 5b 

```{r}
mod9c <- lmer(lnyield ~ lnRR_harvest +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay) #doesn't work with non-correlated intercept and slope
mod9b <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay) #psem needs complete pathways, removing sy without mid or end season density

valid_sy <- unique(mod9c@frame$sy) #just site years in mod2
valid_sy <- unique(mod9b@frame$sy)

newdat <- expand.grid(
  ccbm_sc = seq(min(final_datay$ccbm_sc, na.rm = TRUE),
                max(final_datay$ccbm_sc, na.rm = TRUE),
                length.out = 100),
  DPP_sc = mean(final_datay$DPP_sc, na.rm = TRUE),
  lnRR_harvest = mean(final_datay$lnRR_harvest, na.rm = TRUE), 
  lnRR_4WAP = mean(final_datay$lnRR_4WAP, na.rm = TRUE), 
  sy = valid_sy
)
newdat$pred_sy <- predict(mod9b, newdata = newdat, re.form = NULL)
newdat$pred_pop <- predict(mod9b, newdata = newdat, re.form = NA)



pred_int2 <- predictInterval(
  mod9c,
  newdata = newdat,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat$fit   <- pred_int2$fit
newdat$lwr   <- pred_int2$lwr
newdat$upr   <- pred_int2$upr

figure5b_end <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_sy, color = sy),
            size = 0.75, alpha = 0.4) +
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat,
            aes(x = ccbm_sc, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
  labs(x = "Cover crop biomass",
       y = "Soybean Yield", 
       color = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#7294D4", "#7294D4",   #delaware
                                "#FAD510","#FAD510", #illinois
                                "#C52E19",  #louisiana
                                "#046C9A",  #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A", #missouri
                                "#CEAB07", "#CEAB07", "#CEAB07", #nebraska
                                
                                "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(5, "pt"),
       legend.text = element_text(size = 7),
       legend.title = element_text(size = 8),
       legend.key.size = unit(0.2, "cm"),
       legend.key.spacing.y = unit(0.1, "cm"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE, title.position = "top", title.hjust = 0.5))

figure5b_end

#work out some save details 
ggsave("figure_endseason.png", width = 2.5, height = 2, units = "in", dpi = 300)
```






