---
title: "Cover crop weeds GROW 2026"
author: "L Sencenbaugh"
date: "2026-06-22"
output:
  rmarkdown::html_document:
    theme: hpstr
---

An analytical framework for identifying interaction effects that mediate cover crop surface residue and weed response relationships in IWM studies

Authors: Lilly Sencenbaugh 1, John M. Wallace 1, Eugene Law 2, Muthu Bagavathiannan 3, Kevin Bradley 4, Adam Davis 5, Wesley Everman 6, Michael Flessner 7, Nicholas Jordan 8, Lauren Lazaro 9, John Lindquist 10, Jason Norsworthy 11, Mark VanGessel 12, and Steven B. Mirsky 13

1 Associate Professor and Post-doctoral scholar, Plant Science Department, Pennsylvania State University, University Park, PA, USA; 2 Assistant Professor, Department of Horticulture and Crop Science, Ohio State University, Columbus, OH, USA; 3 Professor, Soil and Crop Sciences, Texas A&M University, College Station, TX; 4 Professor, Division of Plant Sciences, University of Missouri, Columbia, MO, USA; 5 Professor, Department of Crop Sciences, University of Illinois, Urbana, IL, USA; 6 Assistant Professor, Iowa State University, Department of Agronomy, Ames, IA, USA; 7 Professor, Virginia Tech, School of Plant and Environmental Sciences, Blacksburg, VA, USA; 8 Professor, Department of Agronomy and Plant Genetics, University of Minnesota, St. Paul, MN, USA; 9 Assistant Professor, University of Georgia; 10 Professor, University of Nebraska; 11 Professor, University of Arkansas, Department of Crop, Soil and Environmental Sciences, Fayetteville, AK, USA; 12 Professor, University of Delaware, Georgetown, DE, USA; 13 Director of Digital Agriculture, USDA-ARS Sustainable Agricultural Systems Laboratory, Beltsville, MD, USA. 


*Introduction to study*

Managing cover crops to produce high-biomass cover crop surface residues is a promising integrated weed management strategy by reducing survival and fitness of annual weeds. Here we seek to address the context dependence that influences cover crop - weed relationships that impact management recommendations. We used piecewise structural equation models (pSEM) and hierarchical modeling procedures to discern the direction and magnitudes of interactive effects between cover crops, weeds, and soybean yield. 


Required packages

```{r, install packages, include = FALSE}
library(plyr) #data organization
library(dplyr) #data organization
library(ggplot2) #figures
library(lme4) #linear mixed effects models
library(lmerTest) #linear mixed effects models
library(readr) #importing csv
library(tidyr) #data organization
#install.packages("piecewiseSEM")
library(piecewiseSEM) #piecewise structural equation modeling 
library(car) #stats
#install.packages("prettydoc")
library(prettydoc) #aesthetics 
library(influence.ME) #influential points
library(broom.mixed) #random effect figures
library(effects) #effects plots
library(ggpubr) #plot stuff
library(ggeffects) #plot effects
#library(semEff) #sem effects
library(Gmisc) #plot stuff
library(scales) #plot stuff
library(grid) #plot stuff
library(MuMIn) #pairwise and group comparisons
library(lattice) # dot plot
library(merTools) #create confidence intervals from lmer models
library(performance) #checking aic and r2 values
```



#### Import the data

```{r    echo = TRUE, warning = FALSE}
#setwd("~/CCHIT 2025")
cchit <- read_csv("biomass_master.csv") 
cchit <- as.data.frame(cchit)

# setting factors and numbers
cchit$state <- as.factor(cchit$state)
cchit$year <- as.factor(cchit$year)
cchit$plot <- as.factor(cchit$plot)
cchit$term.trmt <- as.factor(cchit$term.trmt)
cchit$herb.trmt <- as.factor(cchit$herb.trmt)
cchit$region <- as.factor(cchit$region)
cchit$DPP <- as.numeric(cchit$DPP) # days pre plant (negative values, more negative means earlier termination before planting)
cchit$sy <- interaction(cchit$state, cchit$year) # site year by combining site*year
cchit$plotID <- as.factor(cchit$plotID)

#str(cchit)

cchit <- cchit %>% 
  dplyr :: rename(ccbm = "cc.bm.kgha", 
         yield = "yield.kgha", 
         weedbmh = "weed.bm.harv.kgha", 
         weedbmt = "weed.bm.term.kgha", 
         densityh = "density.harv.allspp.m2", 
         densityt = "density.term.allspp.m2", 
         sbpop = "soybean.pop.ha") #renaming 

# converting biomass in kilograms to megagrams
cchit$weedbmh_mg <- as.numeric(cchit$weedbmh/1000) 
cchit$weedbmt_mg <- as.numeric(cchit$weedbmt/1000)
cchit$ccbm_mg <- as.numeric(cchit$ccbm/1000)

# 4 weeks

itsbeen4weekssinceyoulookedatme <- read_csv("weed count data.csv") 

wap4 <- as.data.frame(itsbeen4weekssinceyoulookedatme)

# setting factors and numbers
wap4$state <- as.factor(wap4$state)
wap4$year <- as.factor(wap4$year)
wap4$block <- as.factor(wap4$block)
wap4$plot <- as.factor(wap4$plot)
wap4$term_trmt <- as.factor(wap4$term_trmt)
wap4$herb_trmt <- as.factor(wap4$herb_trmt)
wap4$wap <- as.numeric(wap4$wap)  
wap4$sy <- interaction(wap4$state, wap4$year) # site year by combining site*year
wap4$plotID <- as.factor(wap4$plotID)
wap4 <- wap4 %>%
  filter(!wap == "0") ## same thing as at termination--only 1 site year has this
str(wap4)
```

Weed density mid-season

```{r warning = FALSE}
##creating the mid season density df

sum_per <- function(x) {
  if (all(is.na(x))) NA_real_ else sum(x, na.rm = TRUE) # have to remove NAs for the log aspect AND psem gets mad if they're in there
}

quadsize <- function(x) {
  if (all(is.na(x))) NA else first(na.omit(x))
} # some site years have different plot sizes

totals <- wap4 %>% #wap = weeks after planting
  dplyr :: group_by(sy, plotID, wap, term_trmt, herb_trmt) %>%
  dplyr :: summarise(
    total_ct1 = sum_per(seq_ct_1), #frame 1
    total_ct2 = sum_per(seq_ct_2), #frame 2
    total_ct3 = sum_per(seq_qt_3 ), #frame 3
    quad1_size = quadsize(seq_quad_area_1),
    quad2_size = quadsize(seq_quad_area_2),
    quad3_size = quadsize(seq_quad_3),
    total_perm2_1 = sum_per(Count_per1m_1), #normalizing to 1m2
    total_perm2_2 = sum_per(Count_per1m_2),
    total_perm2_3 = sum_per(Count_per1m_3),
    .groups = "drop"
  )

totals <- as.data.frame(totals)

#making data frame long based on quadrat
totals <- totals %>% 
  dplyr :: rename(term.trmt = term_trmt)
totals <- totals %>%
  dplyr :: rename(herb.trmt = herb_trmt)

totals_long <- totals %>%
  pivot_longer(
    cols = c(total_perm2_1, total_perm2_2), #only texas had 3 frames
    names_to = "quadrat",
    values_to = "total_perm2"
  ) %>%
  dplyr :: mutate(
    quadrat = ifelse(quadrat == "total_perm2_1", "quad1", "quad2")
  )

totals_long <- totals_long %>%
  dplyr :: mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

totals_long <- as.data.frame(totals_long)

totals_joined <- totals_long %>% #combine!!!
  left_join(
    cchit %>% 
      dplyr::select(sy, plotID, term.trmt, herb.trmt, DPP, ccbm_mg),
    by = c("sy", "plotID", "term.trmt", "herb.trmt")
  )

totals_weedy <- totals_joined %>% filter(herb.trmt != "Standard") #weedy totals not standard

str(totals_weedy)
```


We are using the log response ratio

```{r warning = FALSE}

group_means <- totals_weedy %>%
  dplyr :: group_by(sy, plotID, term.trmt, DPP, ccbm_mg, wap, quadrat) %>%
  dplyr :: summarise(
    mean_density = mean(total_perm2, na.rm = TRUE),
    .groups = "drop"
  )

ref <- group_means %>% #reference level i.e. the denominator
  dplyr :: filter(term.trmt == "No cover") %>% 
  dplyr :: rename(ref_density = mean_density) %>% 
  dplyr :: select(sy, plotID, wap, quadrat, ref_density) #wap = weeks after plant

group_means <- group_means %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))

rr_density <- group_means %>% #combine!
  left_join(ref, by = c("sy", "block", "wap", "quadrat"), suffix = c("", "_ref")) %>% #added the _ref to tell apart for later
  mutate(
    RR = mean_density / ref_density,
    lnRR = log(RR + 0.000001) #log response ratio
  )


rr_density <- as.data.frame(rr_density)

rr_density <- rr_density %>%
  dplyr :: filter(!(wap < 2), !(wap > 10)) #MID season

str(rr_density) 
```


Density end of season

```{r warning = FALSE}
cchitnh <- subset(cchit, herb.trmt == "Weedy") # subset out herbicide treated
cchitnh <- cchitnh %>%
  dplyr :: mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

group_means <- cchitnh %>%
  dplyr :: group_by(state, year, sy, term.trmt, plotID, DPP, ccbm_mg) %>%
  dplyr :: summarize(
    mean_biomass = mean(densityh, na.rm = TRUE),
    .groups = "drop"
  ) 


ref <- group_means %>% #basically repeating above but with density at harvest!
  dplyr :: filter(term.trmt == "No cover") %>%
  dplyr :: rename(ref_biomass = mean_biomass) %>%
   dplyr :: select(sy, plotID, ref_biomass)

group_means <- group_means %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
 dplyr ::  mutate(block = substr(plotID, 2, 3))



rr_data <- group_means %>%
  left_join(ref, by = "sy", suffix = c("", "_ref")) %>%
  mutate(
    RR = mean_biomass / ref_biomass,
    lnrr1 = log(RR + 0.000001)
  )


rr_data$RR <- as.numeric(rr_data$RR)
rr_data <- as.data.frame(rr_data)
str(rr_data)

rr_data$lnrr1 <- log(rr_data$RR + 0.000001) #log!



# why are we losing site years?

#sum(is.nan(group_means$mean_biomass))
#group_means %>%
 # filter(is.nan(mean_biomass))
#ref %>%
 # filter(is.nan(ref_biomass))

# missing data (density)


```


Making a new dataset with density at harvest and 4 weeks combined 

```{r}
rr_data <- rr_data %>%
  dplyr::rename(lnRR_harvest = lnrr1) #harvest

rr_density <- rr_density %>%
 dplyr:: rename(lnRR_4WAP = lnRR) #mid

rr_data <- rr_data %>%
 dplyr:: mutate(across(c(sy, plotID, block), as.character))

rr_density <- rr_density %>%
  dplyr::mutate(across(c(sy, plotID, block), as.character))

rr_density_mean <- rr_density %>%
dplyr::  group_by(sy, block, DPP, ccbm_mg) %>%
dplyr::  summarise(lnRR_4WAP = mean(lnRR_4WAP, na.rm = TRUE), .groups = "drop")

final_data <- rr_data %>%
  dplyr :: left_join(rr_density_mean, by = c("sy", "block", "DPP", "ccbm_mg"))

str(final_data)
```


Checking for specific outliers
```{r, warning = FALSE}
ggplot(final_data, aes(x = DPP, y = lnRR_4WAP, color = state)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_smooth(method = "lm", fullrange = TRUE, size = 1, alpha = 0.2) +
  theme_light()
```

## Quantity vs Quality

For this PSEM model we have 2 component models

2. weed density 4 weeks after planting ~ cover crop biomass
1. cover crop biomass ~ days pre plant


Scaling DPP and biomass
```{r}
final_data$ccbm_sc <- scale(final_data$ccbm_mg)
final_data$DPP_sc  <- scale(final_data$DPP)
```



```{r}
mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod1a <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc || sy), data = final_data) # uncorrelated random effects
anova(mod1a, mod1) # correlated is better!
summary(mod1)

Anova(mod1, type = "III", test = "Chisq")
```


#### Table 1 

```{r}
final_data1 <- final_data %>%
  filter(!is.na(lnRR_4WAP), !is.infinite(lnRR_4WAP), !is.nan(lnRR_4WAP)) 



mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data1, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data1) # uncorrelated random effects
anova(mod2a, mod2) 


mod2c <- lm(lnRR_4WAP ~ ccbm_sc, data = final_data1 )
summary(mod2c)
AIC(mod2c)
r2(mod2c)

mod2d <- lmer(lnRR_4WAP ~ ccbm_sc + (1 | sy), data = final_data1)
summary(mod2d)
AIC(mod2d)
r2(mod2d)

mod2b <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data1) 
summary(mod2b)
r2(mod2b)
AIC(mod2b)

mod2e <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc || sy), data = final_data1) 
summary(mod2e)
r2(mod2e)
AIC(mod2e)


anova(mod2b, mod2e)

summary(mod2)
VarCorr(mod2)
Anova(mod2, type = "III", test.statistic = "Chisq")
```


## Figure 1



```{r}
state_max <- final_data %>%
  group_by(sy) %>%
  summarise(max_ccbm = max(ccbm_mg, na.rm = TRUE))

highest_common <- min(state_max$max_ccbm)

#use this for the figure I think??
state_year_summary <- final_data %>%
  group_by(state, year) %>%
  summarise(
    mean_ccbm = mean(ccbm_mg, na.rm = TRUE),
    median_ccbm = median(ccbm_mg, na.rm = TRUE),
    max_ccbm = max(ccbm_mg, na.rm = TRUE),
    sd_ccbm = sd(ccbm_mg, na.rm = TRUE),
    n = sum(!is.na(ccbm_mg)),
    se_ccbm = sd_ccbm / sqrt(n),
    .groups = "drop"
  )

#order from low to high by state
state_order <- state_year_summary %>%
  group_by(state) %>%
  summarise(state_mean = mean(mean_ccbm)) %>%
  arrange(state_mean) %>%
  pull(state)

state_year_summary$state <- factor(state_year_summary$state,
                                   levels = state_order)

#removing sy that are not in the model
state_year_summary$sy <- interaction(state_year_summary$state, state_year_summary$year)
#state_year_summary <- subset(state_year_summary, !state_year_summary$state == "TX")
#state_year_summary <- subset(state_year_summary, !(state_year_summary$sy %in% c("DE.2019", "DE.2020", "IL.2017", "MD.2018", "MD.2019", "MD.2020", "MN.2017", "MN.2018", "MN.2020","MO.2018", "NE.2017")))
```



```{r}
#how to make jitter same for the points AND error bars

jit <- position_jitter(width = 0, height = 0.2, seed = 13)


yearmax <- ggplot(state_year_summary,
       aes(x = mean_ccbm, y = state, fill = state, shape = year)) +
  
  scale_shape_manual("Year", values = c(21, 22, 23, 24)) +
  scale_fill_manual("Year", values = c(   "#F3DF6C", "#CEAB07", "#FAD510","#7294D4", "#B40F20","#C52E19","#FBA72A", "#972D15","#F11B00", "#85D4E3", "#046C9A")) + 
  
  geom_errorbar(aes(xmin = mean_ccbm - se_ccbm,
                  xmax = mean_ccbm + se_ccbm),
              width = 0.3,
              position = jit) +
  geom_point(size = 1.7, position = jit) +
  labs(x = expression(Cover~crop~biomass~(Mg~ha^-1)),
       y = "State",
       fill = "state") +
  scale_x_continuous(limits = c(0, 8))+
  
  theme_light() +
  coord_fixed(ratio = 0.75) +
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))


yearmax
#ggsave("yearaverages.png", width = 3.5, height = 3.5, units = "in", dpi = 300)
```


## Figure 2 

```{r}
rd1 <- broom.mixed::tidy(mod1, effects = "ran_vals", conf.int = TRUE, response = TRUE)
rd1 <- as.data.frame(rd1) 
interceptd1 <- subset(rd1, term == "(Intercept)")
sloped1 <- subset(rd1, term == "DPP_sc")
sloped1$slope <- sloped1$estimate
sloped1$slope.low <- sloped1$conf.low
sloped1$slope.high <- sloped1$conf.high
df.biplotd1 <- cbind(interceptd1, sloped1$slope, sloped1$slope.low, sloped1$slope.high)
df.biplotd1 <- as.data.frame(df.biplotd1)

wes_yellows <- c("#FAD510", "#E1BD6D", "#9A8822", "#EBCC2A", "#E1AF00", "#E3B710", "#F2AD00", "#FDD262", "#E2D200", "#F3DF6C", "#CEAB07", "#FAD77B", "#FBA72A")
wes_reds <- c("#A42820", "#9B110E", "#550307", "#CB2314", "#F2300F", "#C93312", "#EF5703", "#F11B00", "#FF0000", "#B40F20", "#972D15", "#FD6467", "#5B1A18", "#BD3027", "#C52E19")
wes_blues <- c("#273046", "#899DA4", "#35274A", "#3B9AB2", "#78B7C5", "#3A9AB2", "#6FB2C1", "#91BAB6", "#5BBCD6", "#046C9A", "#ABDDDE", "#46ACC8", "#85D4E3", "#7294D4", "#90D4CC", "#7FC0C6", "#0A9F9D", "#5785C1")


figure1b <-
  ggplot(df.biplotd1, 
         aes (x = estimate, 
              y = sloped1$slope, xmin = conf.low, 
              xmax = conf.high,  
              ymin = sloped1$slope.low, 
              ymax = sloped1$slope.high, 
              fill = level, color = level)) + 
  geom_hline(yintercept = 0, color = "gray85", size=0.75) +
  geom_vline(xintercept = 0, color = "gray85", size=0.75) +
  geom_errorbar(size=0.5, alpha = 0.3) + 
  geom_errorbarh(size=0.5, alpha = 0.3) + 
  geom_point(size = 1.5, shape = 21, alpha = 1.0, stroke = 0.8) +
  labs(x = "Conditional mode (random intercept)",
       y = "Conditional mode (random slope)", 
       color = "Site-year", fill = "Site-year") + 

  theme_light () + 
  scale_color_manual(values = c("#68180D","#68180D", #ar
                                "#266B7E", "#266B7E", "#266B7E", "#266B7E", #de
                                "#A18902", "#A18902", "#A18902", "#A18902",#il
                                "#68180D", "#68180D", "#68180D", #la
                                "#244D51", "#244D51", "#244D51", #md
                                "#AE8224", "#AE8224","#AE8224","#AE8224", #mn
                                "#AE8224", "#AE8224","#AE8224","#AE8224", #mo
                                "#68180D","#68180D","#68180D", #nc
                                "#9D8206","#9D8206", "#9D8206", "#9D8206", #ne
                                "#3F1309",  "#3F1309","#3F1309","#3F1309", #tx
                                "#266B7E", "#266B7E", "#266B7E", "#266B7E"))+ #va
  scale_fill_manual(values = c("#B40F20","#B40F20",                            #arkansas
                               "#7294D4", "#7294D4", "#7294D4", "#7294D4",    # delaware
                                "#FAD510", "#FAD510", "#FAD510",  "#FAD510",  # illinois
                                "#C52E19", "#C52E19", "#C52E19",               #louisiana
                                "#046C9A", "#046C9A", "#046C9A",              #maryland
                                "#F3DF6C", "#F3DF6C","#F3DF6C","#F3DF6C",     #minnesota
                                "#FBA72A", "#FBA72A","#FBA72A",  "#FBA72A",    #missouri
                                "#F11B00","#F11B00","#F11B00",                 #north carolina
                                "#CEAB07","#CEAB07", "#CEAB07", "#CEAB07",    #nebraska
                                "#972D15", "#972D15", "#972D15", "#972D15",   # texas
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(5, "pt"),
       legend.text = element_text(size = 7),
       legend.title = element_text(size = 8),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 6, byrow = TRUE, title.position = "top", title.hjust = 0.5))
 
figure1b

#save to make it look really pretty
#ggsave("figure1b.png", width = 6, height = 2.5, units = "in", dpi = 300)

```




```{r}
final_data2 <- final_data %>%
  filter(!is.na(DPP_sc)) 

mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data2, na.action = na.omit)
newdat1 <- expand.grid(
  DPP_sc = seq(min(final_data2$DPP_sc, na.rm = TRUE),
                max(final_data2$DPP_sc, na.rm = TRUE),
                length.out = 100),
  sy = final_data2$sy
)

newdat1$pred <- predict(mod1, newdata = newdat1,
                         re.form = NULL,
                         allow.new.levels = TRUE) # had to add this due to the missing site years in the model (that do show up in the full dataframe)

newdat1$pred_pop <- predict(mod1, newdata = newdat1, re.form = NA)
```

```{r warning = FALSE}
pred_int <- predictInterval(
  mod1,
  newdata = newdat1,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat1$fit   <- pred_int$fit
newdat1$lwr   <- pred_int$lwr
newdat1$upr   <- pred_int$upr
```

```{r}
figure1a <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat1,
            aes(x = DPP_sc, y = pred, color = sy),
            size = 0.75, alpha = 0.4) +
  
  # Population-level line (black, bold, on top)
  geom_line(data = newdat1,
            aes(x = DPP_sc, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat1,
            aes(x = DPP_sc, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
 
  scale_color_manual(values = c("#B40F20","#B40F20",                          #arkansas
                               "#7294D4", "#7294D4", "#7294D4", "#7294D4",    # delaware
                                "#FAD510", "#FAD510", "#FAD510",  "#FAD510",  # illinois
                                "#C52E19", "#C52E19", "#C52E19",               #louisiana
                                "#046C9A", "#046C9A", "#046C9A",              #maryland
                                "#F3DF6C", "#F3DF6C","#F3DF6C","#F3DF6C",     #minnesota
                                "#FBA72A", "#FBA72A","#FBA72A",  "#FBA72A",    #missouri
                                "#F11B00","#F11B00","#F11B00",                 #north carolina
                                "#CEAB07","#CEAB07", "#CEAB07", "#CEAB07",    #nebraska
                                "#972D15", "#972D15", "#972D15", "#972D15",   # texas
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
  
  
  labs(
    y = "CCSR biomass", 
    x = "Termination timing", 
       color = "Site-year") + 
  theme_light()+
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none")


figure1a

#save to make it look really pretty
#ggsave("figure1a.png", width = 6, height = 6, units = "in", dpi = 300)

figure1 <- ggarrange(figure1a, figure1b)
figure1


#ggsave("figure1.png", width = 6, height = 2.5, units = "in", dpi = 300)
```

## lilly check this

Resetting everything without the yucky site years

Weed density mid-season

```{r warning = FALSE}
group_means <- totals_weedy %>%
  dplyr :: group_by(sy, plotID, term.trmt, DPP, ccbm_mg, wap, quadrat) %>%
  dplyr :: summarise(
    mean_density = mean(total_perm2, na.rm = TRUE),
    .groups = "drop"
  )

ref <- group_means %>% 
  dplyr :: filter(term.trmt == "No cover") %>% 
  dplyr :: rename(ref_density = mean_density) %>% 
  dplyr :: select(sy, plotID, wap, quadrat, ref_density)

group_means <- group_means %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
  dplyr :: mutate(block = substr(plotID, 2, 3))

rr_density <- group_means %>%
  dplyr :: left_join(ref, by = c("sy", "block", "wap", "quadrat"), suffix = c("", "_ref")) %>%
  dplyr :: filter(!is.na(ref_density), !is.na(mean_density)) %>%
  dplyr :: mutate(
    RR = ifelse(ref_density == 0, NA, mean_density / ref_density),
    lnRR = log(RR + 0.000001) #psem needs help interpretting log (can't do it within psem code must be done prior)
  ) %>%
  dplyr::select(-plotID_ref) %>% # keep the main columns tidy 
  filter(!is.na(lnRR), !is.infinite(lnRR), !is.nan(lnRR)) 

rr_density <- as.data.frame(rr_density)



str(rr_density)
```
Density at harvest

```{r warning = FALSE}
cchitnh <- subset(cchit, herb.trmt == "Weedy") # subset out herbicide treated
cchitnh <- cchitnh %>%
  dplyr::mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

group_means <- cchitnh %>%
  dplyr::group_by(state, year, sy, term.trmt, plotID, DPP, ccbm_mg) %>%
  dplyr::summarize(
    mean_biomass = mean(densityh, na.rm = TRUE),
    .groups = "drop"
  ) 

ref <- group_means %>%
  dplyr::filter(term.trmt == "No cover") %>%
  dplyr::rename(ref_biomass = mean_biomass) %>%
   dplyr :: select(sy, plotID, ref_biomass)

group_means <- group_means %>%
  dplyr::mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
  dplyr::mutate(block = substr(plotID, 2, 3))

rr_data <- group_means %>%
  dplyr::left_join(ref, by = "sy", suffix = c("", "_ref")) %>%
  dplyr::filter(!is.na(ref_biomass), !is.na(mean_biomass)) %>%
  dplyr::mutate(RR = mean_biomass / ref_biomass) %>%
  dplyr::select(-plotID_ref, -block_ref) 

rr_data[sapply(rr_data, is.infinite)] <- NA
rr_data <- na.omit(rr_data)


rr_data$RR <- as.numeric(rr_data$RR)
rr_data <- as.data.frame(rr_data)
str(rr_data)

rr_data$lnrr <- log(rr_data$RR + 0.000001) 
```

Making a new dataset with density at harvest and 4 weeks combined 

```{r}
rr_data <- rr_data %>%
  dplyr::rename(lnRR_harvest = lnrr)

rr_density <- rr_density %>%
 dplyr:: rename(lnRR_4WAP = lnRR)

rr_data <- rr_data %>%
 dplyr:: mutate(across(c(sy, plotID, block), as.character))

rr_density <- rr_density %>%
  dplyr::mutate(across(c(sy, plotID, block), as.character))

rr_density_mean <- rr_density %>%
dplyr::  group_by(sy, block, DPP, ccbm_mg) %>%
dplyr::  summarise(lnRR_4WAP = mean(lnRR_4WAP, na.rm = TRUE), .groups = "drop")

final_data <- rr_data %>%
  dplyr :: left_join(rr_density_mean, by = c("sy", "block", "DPP", "ccbm_mg"))

final_data$sy <- as.factor(final_data$sy)
str(final_data)
```


Removing specific outliers
```{r, warning = FALSE}
#x11()
#plot(final_data$DPP, final_data$lnRR_4WAP, pch=19, cex=2, col="goldenrod")

# identify points
#idx <- identify(final_data$DPP, final_data$lnRR_4WAP, labels = final_data$plotID)

# confirm
#final_data[idx, c("plotID", "DPP", "lnRR_harvest")]

# remove by row

final_data <- subset(final_data, !(final_data$plotID %in% c("M73MW", "D73LW", "V02LW")))

ggplot(final_data, aes(x = DPP, y = lnRR_4WAP, color = state)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_smooth(method = "lm", fullrange = TRUE, size = 1, alpha = 0.2) +
  theme_light()
```

```{r}
final_data$ccbm_sc <- scale(final_data$ccbm_mg)
final_data$DPP_sc  <- scale(final_data$DPP)
```

### Model 1 (cover crop biomass ~ days pre planting)


```{r}
mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod1a <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc || sy), data = final_data) # uncorrelated random effects
anova(mod1a, mod1) # correlated is better!
summary(mod1)
Anova(mod1, type = "III", test = "Chisq")
```


```{r   fig.width = 5, fig.height = 5}
residuals <- resid(mod1)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod1, type="link"), pearson = residuals(mod1, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod1, vars = "Cook")
dotplot(ranef(mod1))
```

```{r, fig.height = 15, fig.width = 10}
re <- ranef(mod1)$sy
re_df <- re %>%
  tibble::rownames_to_column("sy") %>%
  rename(rand_intercept = `(Intercept)`)
fix_int <- fixef(mod1)[1]
re_df <- re_df %>%
  mutate(intercept = fix_int + rand_intercept)
final_data1 <- final_data %>%
  left_join(re_df, by = "sy")
final_data1 <- final_data1 %>%
  mutate(sy = reorder(sy, intercept))

ggplot(final_data1, aes(x = DPP_sc.x, y = ccbm_sc)) +
  geom_point(size = 2, alpha = 0.2, pch = 16) +
  geom_smooth(method = "lm", se = TRUE,
              fullrange = TRUE, 
              size = 0.75, alpha = 0.5,
              col = "hotpink") +
  theme_light() + 
  theme(axis.title = element_text(size = 14),
        legend.title = element_text(size = 14), 
        aspect.ratio = 1.5)+
  labs(y = "cover crop biomass scaled", x = "days pre planting")+
  facet_wrap(~sy)
```



### Model 2 (lnRR ~ ccbm)


```{r}

mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated




summary(mod2)
VarCorr(mod2)
Anova(mod2, type = "III", test.statistic = "Chisq")
```


```{r}

mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data) # uncorrelated random effects
anova(mod2a, mod2) # correlated is better!



summary(mod2)
VarCorr(mod2)
Anova(mod2, type = "III", test.statistic = "Chisq")
```

```{r   fig.width = 5, fig.height = 5}
residuals <- resid(mod2)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod2, type="link"), pearson = residuals(mod2, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod2, vars = "Cook")
dotplot(ranef(mod2))
```

```{r, fig.height = 15, fig.width = 10}
re <- ranef(mod2)$sy
re_df <- re %>%
  tibble::rownames_to_column("sy") %>%
  rename(rand_intercept = `(Intercept)`)
fix_int <- fixef(mod2)[1]
re_df <- re_df %>%
  mutate(intercept = fix_int + rand_intercept)
final_data1 <- final_data %>%
  left_join(re_df, by = "sy")
final_data1 <- final_data1 %>%
  mutate(sy = reorder(sy, intercept))

ggplot(final_data1, aes(x = ccbm_sc.x, y = lnRR_4WAP)) +
  geom_point(size = 2, alpha = 0.2, pch = 16) +
  geom_smooth(method = "lm", se = TRUE,
              fullrange = TRUE, 
              size = 0.75, alpha = 0.5,
              col = "seagreen1") +
  theme_light() + 
  theme(axis.title = element_text(size = 14),
        legend.title = element_text(size = 14), 
        aspect.ratio = 1.5)+
  labs(y = "lnRR_4WAP", x = "ccbmscaled")+
  facet_wrap(~sy)
```



### solving for % change in density

First back transforming from the scaled version


MAKE A TABLE with this information
Stick to 14 days pre-plant
make a few thresholds like 50%, 75%, and 90%

Can we isolate a maximum and minimum based on the site year random effect?

```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)
# coefficients 
b0 <- fixef(mod2)[1]
b1 <- fixef(mod2)["ccbm_sc"]
b2 <- fixef(mod2)["DPP_sc"]

# means and SDs
mean_ccbm <- mean(final_data$ccbm_mg, na.rm = TRUE) # MUST REMOVE NAs
sd_ccbm   <- sd(final_data$ccbm_mg, na.rm = TRUE)

mean_dpp <- mean(final_data$DPP, na.rm = TRUE)
sd_dpp   <- sd(final_data$DPP, na.rm = TRUE)


dpp_val <- -14  # termination two weeks before planting 

#scaled to normal
dpp_sc <- (dpp_val - mean_dpp) / sd_dpp

# target level-- so here it's 90% reduction
# lnRR = ln (cover crop treatments/ no cover crops)
target_lnRR <- log(0.20)

biomass_needed <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)
```
This is the version from before the scaling 
I'm checking to see if the number is close
```{r}
mod2 <-  lmer(lnRR_4WAP ~ ccbm_mg + DPP  + (ccbm_mg | sy), data = final_data, na.action = na.omit)
b0 <- fixef(mod2)[1]
b1 <- fixef(mod2)["ccbm_mg"]
b2 <- fixef(mod2)["DPP"]
dpp_val <- -14 # termination right before planting
target_lnRR <- log(0.1) # 90% decrease in density 
biomass_needed <- (target_lnRR - b0 - b2 * dpp_val) / b1
```

It is so close I think we call it the same. This just confirms the back transforming is actually working! 


Different termination timings for 90% decline in density (citation--I think liebman)

```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)
b0 <- fixef(mod2)[1]
b1 <- fixef(mod2)["ccbm_sc"]
b2 <- fixef(mod2)["DPP_sc"]
mean_ccbm <- mean(final_data$ccbm_mg, na.rm = TRUE) # MUST REMOVE NAs
sd_ccbm   <- sd(final_data$ccbm_mg, na.rm = TRUE)
mean_dpp <- mean(final_data$DPP, na.rm = TRUE)
sd_dpp   <- sd(final_data$DPP, na.rm = TRUE)
dpp_sc <- (dpp_val - mean_dpp) / sd_dpp
target_lnRR <- log(0.10)

dpp_val <- 0 # termination day of
dpp_sc <- (dpp_val - mean_dpp) / sd_dpp
biomass_needed <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)

dpp_val <- -14 # termination late
dpp_sc <- (dpp_val - mean_dpp) / sd_dpp
biomass_needed <-  mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)

dpp_val <- -28 # termination mid
dpp_sc <- (dpp_val - mean_dpp) / sd_dpp
biomass_needed <-  mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)

dpp_val <- -42 # termination mid
dpp_sc <- (dpp_val - mean_dpp) / sd_dpp
biomass_needed <-  mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)

dpp_val <- -56 # termination early
dpp_sc <- (dpp_val - mean_dpp) / sd_dpp
biomass_needed <-  mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)

```

Termination the day of planting: To cause a 90% reduction in weed density, you need 4.8Mg per hectare. 

Termination 2 weeks before planting: To cause a 90% reduction in weed density, you need 5.5Mg per hectare. 

Termination 4 weeks before planting: To cause a 90% reduction in weed density, you need 6.3 Mg per hectare. 

Termination 6 weeks before planting: To cause a 90% reduction in weed density, you need 7.0 Mg per hectare. 

Termination 8 weeks before planting: To cause a 90% reduction in weed density, you need 7.8 Mg per hectare. 

Nichols et al (2020) didn't have a specific number in text for a 90% reduction in weed density-- they said that none of their sites could realistically achieve a biomass to cause this reduction. I think our numbers align well with their findings. 


Find the max and min and maybe make a figure?? Table I made in word


I'm starting over to be not confused 
```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit) # the model

# coefficients 
b0 <- fixef(mod2)[1] #intercept
b1 <- fixef(mod2)["ccbm_sc"]
b2 <- fixef(mod2)["DPP_sc"]

# means and SDs
mean_ccbm <- mean(final_data$ccbm_mg, na.rm = TRUE) # MUST REMOVE NAs
sd_ccbm   <- sd(final_data$ccbm_mg, na.rm = TRUE)

mean_dpp <- mean(final_data$DPP, na.rm = TRUE)
sd_dpp   <- sd(final_data$DPP, na.rm = TRUE)

#fix the term timing
dpp_val <- -14  # termination two weeks before planting 

#scaled to normal
dpp_sc <- (dpp_val - mean_dpp) / sd_dpp

# target level--  here it's 90% reduction
# lnRR = ln (cover crop treatments/ no cover crops)
target_lnRR <- log(0.10)

biomass_needed <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)
re <- ranef(mod2)$sy #random effect

#slope deviation
slope_dev <- re$ccbm_sc

min_slope <- min(slope_dev)
max_slope <- max(slope_dev)


#slope adjustments
b1_min <- b1 + min_slope
b1_max <- b1 + max_slope

#this is the actual number not the range
biomass_90_pop <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)

#range
#max (you put the min in the denom but it's the max I swear)
biomass_90_max <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1_min)
#min
biomass_90_min <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1_max)



#repeat above for 75%
target_lnRR <- log(0.25)
biomass_needed <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)
re <- ranef(mod2)$sy
slope_dev <- re$ccbm_sc
min_slope <- min(slope_dev)
max_slope <- max(slope_dev)
b1_min <- b1 + min_slope
b1_max <- b1 + max_slope
biomass_75_pop <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)
biomass_75_max <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1_min)
biomass_75_min <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1_max)




# 50% reduction
target_lnRR <- log(0.5)
biomass_needed <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)
re <- ranef(mod2)$sy
slope_dev <- re$ccbm_sc
min_slope <- min(slope_dev)
max_slope <- max(slope_dev)
b1_min <- b1 + min_slope
b1_max <- b1 + max_slope
biomass_50_pop <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1)
biomass_50_max <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1_min)
biomass_50_min <- mean_ccbm + 
  sd_ccbm * ((target_lnRR - b0 - b2 * dpp_sc) / b1_max)
```



Holding at 2 Mg/ha 
model would predict x reduction in weed density at 4 wap

Inspection of population-level fixed effect estimates (ß0, ß1) from candidate models and back-transformed predictions of weed suppression levels demonstrate how the model fitting procedure can influence management inferences. At a 2 Mg ha-1 of CCSR biomass, simple linear regression (M1) would predict a XX% reduction in weed density at 4 WAP, whereas a random intercept (M2) and random intercept and slope (M3) model would predict a XX% and XX% reduction in weed density. 

```{r}
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))
final_data$DPP_sc  <- as.numeric(scale(final_data$DPP))
mod <- lm(lnRR_4WAP ~ ccbm_sc , data= final_data, na.action = na.omit)
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc || sy), data = final_data) # uncorrelated random effects



#first make 2Mg in the scaled scale
ccbm_mean <- mean(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sd   <- sd(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sc_2mg <- (2 - ccbm_mean) / ccbm_sd #finding 2 but scaled

#need to pick a reasonable termination time
DPP_mean <- mean(final_data$DPP, na.rm = TRUE)
DPP_sd   <- sd(final_data$DPP, na.rm = TRUE)

DPP_sc_14 <- (-14 - DPP_mean) / DPP_sd #two weeks

newdat <- data.frame(
  ccbm_sc = ccbm_sc_2mg,
  DPP_sc = DPP_sc_14
)

pred_lnRR <- predict(mod, newdata = newdat) #predict
RR <- exp(pred_lnRR) #back transform the lnrr 
percent_reduction <- (1 - RR) * 100 #percent reduction

# mixed model just intercept

mod2c <- lmer(lnRR_4WAP ~ ccbm_sc  + ( 1 | sy), data = final_data) # uncorrelated random effects

pred_lnRR_int <- predict(mod2c, newdata = newdat, re.form = NA)
RR_int <- exp(pred_lnRR_int) #back transform the lnrr 
percent_reduction_int <- (1 - RR_int) * 100 #percent reduction



#correlated mixed model
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))
final_data$DPP_sc  <- as.numeric(scale(final_data$DPP))
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc +  (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated


pred_lnRR_corr <- predict(mod2, newdata = newdat, re.form = NA)
RR_corr <- exp(pred_lnRR_corr) #back transform the lnrr 
percent_reduction_corr <- (1 - RR_corr) * 100 #percent reduction



```



1Mg

```{r}
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))
final_data$DPP_sc  <- as.numeric(scale(final_data$DPP))
mod <- lm(lnRR_4WAP ~ ccbm_sc , data= final_data, na.action = na.omit)
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc || sy), data = final_data) # uncorrelated random effects



#first make 1Mg in the scaled scale
ccbm_mean <- mean(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sd   <- sd(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sc_1mg <- (1 - ccbm_mean) / ccbm_sd #finding 2 but scaled

newdat <- data.frame(
  ccbm_sc = ccbm_sc_1mg
 
)

pred_lnRR <- predict(mod, newdata = newdat) #predict
RR <- exp(pred_lnRR) #back transform the lnrr 
percent_reduction <- (1 - RR) * 100 #percent reduction

# mixed model just intercept

mod2c <- lmer(lnRR_4WAP ~ ccbm_sc  + ( 1 | sy), data = final_data) # uncorrelated random effects

pred_lnRR_int <- predict(mod2c, newdata = newdat, re.form = NA)
RR_int <- exp(pred_lnRR_int) #back transform the lnrr 
percent_reduction_int <- (1 - RR_int) * 100 #percent reduction



#correlated mixed model
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))
final_data$DPP_sc  <- as.numeric(scale(final_data$DPP))
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc +  (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated


pred_lnRR_corr <- predict(mod2, newdata = newdat, re.form = NA)
RR_corr <- exp(pred_lnRR_corr) #back transform the lnrr 
percent_reduction_corr <- (1 - RR_corr) * 100 #percent reduction



```



4Mg

```{r}
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))
final_data$DPP_sc  <- as.numeric(scale(final_data$DPP))
mod <- lm(lnRR_4WAP ~ ccbm_sc , data= final_data, na.action = na.omit)
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc  + (ccbm_sc || sy), data = final_data) # uncorrelated random effects



#first make 4Mg in the scaled scale
ccbm_mean <- mean(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sd   <- sd(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sc_4mg <- (4 - ccbm_mean) / ccbm_sd #finding 2 but scaled


newdat <- data.frame(
  ccbm_sc = ccbm_sc_4mg
)

pred_lnRR <- predict(mod, newdata = newdat) #predict
RR <- exp(pred_lnRR) #back transform the lnrr 
percent_reduction <- (1 - RR) * 100 #percent reduction

# mixed model just intercept

mod2c <- lmer(lnRR_4WAP ~ ccbm_sc  + ( 1 | sy), data = final_data) # uncorrelated random effects

pred_lnRR_int <- predict(mod2c, newdata = newdat, re.form = NA)
RR_int <- exp(pred_lnRR_int) #back transform the lnrr 
percent_reduction_int <- (1 - RR_int) * 100 #percent reduction



#correlated mixed model
final_data$ccbm_sc <- as.numeric(scale(final_data$ccbm_mg))
final_data$DPP_sc  <- as.numeric(scale(final_data$DPP))
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc +  (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated


pred_lnRR_corr <- predict(mod2, newdata = newdat, re.form = NA)
RR_corr <- exp(pred_lnRR_corr) #back transform the lnrr 
percent_reduction_corr <- (1 - RR_corr) * 100 #percent reduction



```



### PSEM

Starting-- most basic, no end-season density

DPP -> ccbm 
ccbm -> lnrr_mid
```{r}
psemDensity <- psem(
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity, .progressBar = FALSE)
# p = 0.059
fisherC(psemDensity)
#Fisher's: 5.643 on 2 df, p = 0.06	
summary(psemDensity)
```

The p = 0.059 is stressing me out because it is so borderline

DPP -> ccbm 
ccbm -> lnrr_mid
```{r}
psemDensity1 <- psem(
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity1, .progressBar = FALSE)
# saturated
fisherC(psemDensity1)
#Fisher's: 5.643 on 2 df, p = 0.06	
summary(psemDensity1)
```


Confirmed-- we are keeping this as the final model


## Figure 2 ig

```{r}
mod2b <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data) # remove dpp
summary(mod2b)
r2(mod2b)
AIC(mod2b)
```


```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data) # remove dpp
rd2 <- broom.mixed::tidy(mod2, effects = "ran_vals", conf.int = TRUE, response = TRUE)
rd2 <- as.data.frame(rd2) 
interceptd2 <- subset(rd2, term == "(Intercept)")
sloped2 <- subset(rd2, term == "ccbm_sc")
sloped2$slope <- sloped2$estimate
sloped2$slope.low <- sloped2$conf.low
sloped2$slope.high <- sloped2$conf.high
df.biplotd2 <- cbind(interceptd2, sloped2$slope, sloped2$slope.low, sloped2$slope.high)
df.biplotd2 <- as.data.frame(df.biplotd2)

figure1d <-
  ggplot(df.biplotd2, 
         aes (x = estimate, 
              y = sloped2$slope, xmin = conf.low, 
              xmax = conf.high,  
              ymin = sloped2$slope.low, 
              ymax = sloped2$slope.high, 
              fill = level, color = level)) + 
  geom_hline(yintercept = 0, color = "gray85", size=0.75) +
  geom_vline(xintercept = 0, color = "gray85", size=0.75) +
  geom_errorbar(size=0.5, alpha = 0.3) + 
  geom_errorbarh(size=0.5, alpha = 0.3) + 
  geom_point(size = 1.5, shape = 21, alpha = 1.0, stroke = 0.8) +
  labs(x = "Conditional mode (random intercept)",
       y = "Conditional mode (random slope)", 
       color = "Site-year", fill = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#266B7E", "#266B7E",  
                                "#A18902", "#A18902", 
                                "#68180D", "#68180D", 
                                "#023146", 
                                "#AE8224",
                                "#AE8224", "#AE8224",
                                "#9D8206", "#9D8206","#9D8206",  
                                "#266B7E", "#266B7E", "#266B7E", "#266B7E"))+

   scale_fill_manual(values = c("#7294D4", "#7294D4",  #delaware
                                "#FAD510", "#FAD510", #illinois
                                "#C52E19", "#C52E19", #louisiana
                                "#046C9A", #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A",#missouri
                                "#CEAB07", "#CEAB07", "#CEAB07", #nebraska
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) +
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))




figure1d
#ggsave("figure1d.png", width = 7, height = 2.5, units = "in", dpi = 300)
```



```{r}

mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data) # remove dpp

# am having issues with dropped site years sneaking back into df
# in the dataframe we have lots of site years but in mod2 only 17 have enough observations

valid_sy <- unique(mod2@frame$sy) #just site years in mod2

newdat <- expand.grid(
  ccbm_sc = seq(min(final_data$ccbm_sc, na.rm = TRUE),
                max(final_data$ccbm_sc, na.rm = TRUE),
                length.out = 100),
  DPP_sc = mean(final_data$DPP_sc, na.rm = TRUE),
  sy = valid_sy
)
newdat$pred_sy <- predict(mod2, newdata = newdat, re.form = NULL)

newdat$pred_pop <- predict(mod2, newdata = newdat, re.form = NA)
```


Confidence interval

```{r}
pred_int1 <- predictInterval(
  mod2,
  newdata = newdat,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat$fit   <- pred_int1$fit
newdat$lwr   <- pred_int1$lwr
newdat$upr   <- pred_int1$upr
```


```{r}
figure1c <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_sy, color = sy),
            size = 0.75, alpha = 0.4) +
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat,
            aes(x = ccbm_sc, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
  labs(x = "CCSR biomass",
       y = "Weed density (LRR; 4 WAP)", 
       color = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#7294D4", "#7294D4", #delaware
                                "#FAD510", "#FAD510", #illinois
                                "#C52E19", "#C52E19", #louisiana
                                "#046C9A", #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A",#missouri
                                "#CEAB07", "#CEAB07", "#CEAB07", #nebraska
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
 
   
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
        legend.position = "none")





figure1bottom <- ggarrange(figure1c, figure1d)
figure1bottom

#work out some save details 
ggsave("figure1bottom.png", width = 6, height = 2.5, units = "in", dpi = 300)
```




### Figures

```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)
rd2 <- broom.mixed::tidy(mod2, effects = "ran_vals", conf.int = TRUE, response = TRUE)
rd2 <- as.data.frame(rd2) 
interceptd2 <- subset(rd2, term == "(Intercept)")
sloped2 <- subset(rd2, term == "ccbm_sc")
sloped2$slope <- sloped2$estimate
sloped2$slope.low <- sloped2$conf.low
sloped2$slope.high <- sloped2$conf.high
df.biplotd2 <- cbind(interceptd2, sloped2$slope, sloped2$slope.low, sloped2$slope.high)
df.biplotd2 <- as.data.frame(df.biplotd2)

figure1d <-
  ggplot(df.biplotd2, 
         aes (x = estimate, 
              y = sloped2$slope, xmin = conf.low, 
              xmax = conf.high,  
              ymin = sloped2$slope.low, 
              ymax = sloped2$slope.high, 
              fill = level, color = level)) + 
  geom_hline(yintercept = 0, color = "gray85", size=0.75) +
  geom_vline(xintercept = 0, color = "gray85", size=0.75) +
  geom_errorbar(size=0.5, alpha = 0.3) + 
  geom_errorbarh(size=0.5, alpha = 0.3) + 
  geom_point(size = 1.5, shape = 21, alpha = 1.0, stroke = 0.8) +
  labs(x = "Conditional mode (random intercept)",
       y = "Conditional mode (random slope)", 
       color = "Site-year", fill = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#266B7E", "#266B7E",  
                                "#A18902", "#A18902", 
                                "#68180D", "#68180D", 
                                "#023146", 
                                "#AE8224",
                                "#AE8224", "#AE8224",
                                "#9D8206", "#9D8206","#9D8206",  
                                "#266B7E", "#266B7E", "#266B7E", "#266B7E"))+

   scale_fill_manual(values = c("#7294D4", "#7294D4",  #delaware
                                "#FAD510", "#FAD510", #illinois
                                "#C52E19", "#C52E19", #louisiana
                                "#046C9A", #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A",#missouri
                                "#CEAB07", "#CEAB07", "#CEAB07", #nebraska
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) +
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))




figure1d
#ggsave("figure1d.png", width = 7, height = 2.5, units = "in", dpi = 300)
```




```{r}

mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)

# am having issues with dropped site years sneaking back into df
# in the dataframe we have lots of site years but in mod2 only 17 have enough observations

valid_sy <- unique(mod2@frame$sy) #just site years in mod2

newdat <- expand.grid(
  ccbm_sc = seq(min(final_data$ccbm_sc, na.rm = TRUE),
                max(final_data$ccbm_sc, na.rm = TRUE),
                length.out = 100),
  DPP_sc = mean(final_data$DPP_sc, na.rm = TRUE),
  sy = valid_sy
)
newdat$pred_sy <- predict(mod2, newdata = newdat, re.form = NULL)

newdat$pred_pop <- predict(mod2, newdata = newdat, re.form = NA)
```


Confidence interval

```{r}
pred_int1 <- predictInterval(
  mod2,
  newdata = newdat,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat$fit   <- pred_int1$fit
newdat$lwr   <- pred_int1$lwr
newdat$upr   <- pred_int1$upr
```


```{r}
figure1c <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_sy, color = sy),
            size = 0.75, alpha = 0.4) +
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat,
            aes(x = ccbm_sc, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
  labs(x = "CCSR biomass",
       y = "Weed density (LRR; 4 WAP)", 
       color = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#7294D4", "#7294D4", #delaware
                                "#FAD510", "#FAD510", #illinois
                                "#C52E19", "#C52E19", #louisiana
                                "#046C9A", #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A",#missouri
                                "#CEAB07", "#CEAB07", "#CEAB07", #nebraska
                                "#85D4E3", "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
 
   
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
        legend.position = "none")





figure1bottom <- ggarrange(figure1c, figure1d)
figure1bottom

#work out some save details 
ggsave("figure1bottom.png", width = 6, height = 2.5, units = "in", dpi = 300)
```


```{r}
state_max <- final_data1 %>%
  group_by(sy) %>%
  summarise(max_ccbm = max(ccbm_mg, na.rm = TRUE))

highest_common <- min(state_max$max_ccbm)

#use this for the figure I think??
state_year_summary <- final_data1 %>%
  group_by(state, year) %>%
  summarise(
    mean_ccbm = mean(ccbm_mg, na.rm = TRUE),
    median_ccbm = median(ccbm_mg, na.rm = TRUE),
    max_ccbm = max(ccbm_mg, na.rm = TRUE),
    sd_ccbm = sd(ccbm_mg, na.rm = TRUE),
    n = sum(!is.na(ccbm_mg)),
    se_ccbm = sd_ccbm / sqrt(n),
    .groups = "drop"
  )

#order from low to high by state
state_order <- state_year_summary %>%
  group_by(state) %>%
  summarise(state_mean = mean(mean_ccbm)) %>%
  arrange(state_mean) %>%
  pull(state)

state_year_summary$state <- factor(state_year_summary$state,
                                   levels = state_order)

#removing sy that are not in the model
state_year_summary$sy <- interaction(state_year_summary$state, state_year_summary$year)
#state_year_summary <- subset(state_year_summary, !state_year_summary$state == "TX")
#state_year_summary <- subset(state_year_summary, !(state_year_summary$sy %in% c("DE.2019", "DE.2020", "IL.2017", "MD.2018", "MD.2019", "MD.2020", "MN.2017", "MN.2018", "MN.2020","MO.2018", "NE.2017")))
```



```{r}
#how to make jitter same for the points AND error bars

jit <- position_jitter(width = 0, height = 0.2, seed = 13)


yearmax <- ggplot(state_year_summary,
       aes(x = mean_ccbm, y = state, fill = state, shape = year)) +
  
  scale_shape_manual("Year", values = c(21, 22, 23, 24)) +
#  scale_fill_manual("Year", values = c( "#972D15", "#C52E19", "#F3DF6C", "#CEAB07", "#FAD510","#7294D4", "#FBA72A",  "#85D4E3", "#046C9A")) + 
  
  geom_errorbar(aes(xmin = mean_ccbm - se_ccbm,
                  xmax = mean_ccbm + se_ccbm),
              width = 0.3,
              position = jit) +
  geom_point(size = 1.7, position = jit) +
  labs(x = expression(Cover~crop~biomass~(Mg~ha^-1)),
       y = "State",
       fill = "state") +
  scale_x_continuous(limits = c(0, 8))+
  
  theme_light() +
  coord_fixed(ratio = 0.75) +
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))

#ggsave("yearaverages.png", width = 3.5, height = 3.5, units = "in", dpi = 300)
```




### Difference figure 2 Mg
```{r}
#the scaling
#final_data$ccbm_sc <- scale(final_data$ccbm_mg)
#final_data$DPP_sc  <- scale(final_data$DPP)
#the model
#mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)

ccbm_mean <- mean(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sd   <- sd(final_data$ccbm_mg, na.rm = TRUE)

DPP_mean <- mean(final_data$DPP, na.rm = TRUE)
DPP_sd   <- sd(final_data$DPP, na.rm = TRUE)

ccbm_vals <- c(2) #common highest mean among sites yikes
ccbm_sc_vals <- (ccbm_vals - ccbm_mean) / ccbm_sd #un-scale 

DPP_val <- -14 #focus on 2 weeks pre planting
DPP_sc_val <- (DPP_val - DPP_mean) / DPP_sd #un-scale

model_sy <- unique(mod2@frame$sy) #site year

newdata <- expand.grid(
  sy = model_sy,
  ccbm_sc = ccbm_sc_vals,
  DPP_sc = DPP_sc_val
)

newdata$biomass <- rep(c("2 Mg/ha"), #yikes
                       each = length(model_sy))

#check we didn't remove anything by accident
#length(unique(mod2@frame$sy)) #17 sy
#length(unique(newdata$sy))# 17 sy


```



```{r}
newdata$fit_sy <- predict(mod2, newdata = newdata, re.form = NULL)
newdata$fit_pop <- predict(mod2, newdata = newdata, re.form = NA)
pred_pop <- predictInterval(
  mod2,
  newdata = newdata,
  level = 0.95,
  n.sims = 2000,
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"
)
# convert to %
newdata$percent_sy  <- (1 - exp(newdata$fit_sy)) * 100
newdata$percent_pop <- (1 - exp(newdata$fit_pop)) * 100

newdata$pop_lwr <- (1 - exp(pred_pop$upr)) * 100
newdata$pop_upr <- (1 - exp(pred_pop$lwr)) * 100

pop_data <- data.frame(
  sy = "Population",
  percent_pop = unique(newdata$percent_pop)[1],
  pop_lwr = unique(newdata$pop_lwr)[1],
  pop_upr = unique(newdata$pop_upr)[1]
)

pop_mean <- unique(newdata$percent_pop)[1]

plot_data <- newdata %>%
  mutate(
    deviation = percent_sy - pop_mean
  )

plot_data <- plot_data %>%
  arrange(percent_sy)

plot_data$sy <- factor(plot_data$sy, levels = plot_data$sy)
```


```{r}
figure2b1 <- ggplot(plot_data, aes(y = sy, fill = sy, color = sy)) +
 geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "gray50",
             size = 0.25) + 
  # segment from population mean to site-year
  geom_segment(aes(x = pop_mean, xend = percent_sy,
                   yend = sy),
               color = "gray20", size = 0.5) +
  
  
  # population reference line
  geom_vline(xintercept = pop_mean,
             linetype = "longdash",
             color = "black", 
             size = 0.5) +
  # population reference point
 geom_point(data = pop_data,
           aes(x = percent_pop, y = sy),
           shape = 21,
           size = 3,
           fill = "black",
           color = "black") + 
  # population reference point error bar
geom_errorbarh(data = pop_data,
               aes(xmin = pop_lwr,
                   xmax = pop_upr,
                   y = sy),
               height = 0.5,
               color = "black",
               size = 0.5) +

    # site-year points
  geom_point(aes(x = percent_sy),
             shape = 21, size = 1.25, stroke = 0.7) +
  
    
  scale_color_manual(values = c("#9D8206", #NE
                                "#AE8224",#MN
                                "#AE8224",#MO
                                "#A18902",#IL
                                "#266B7E",#VA
                                "#023146", #DE
                                "#A18902", #IL
                                "#266B7E", #VA
                                "#266B7E",#VA
                                "#023146",#MD
                                "#9D8206",#NE
                                "#9D8206",#NE
                                "#AE8224",#MO
                                "#266B7E",#VA
                                "#68180D", #LA
                                "#023146",#DE
                                "#68180D", #LA
                                "black"))+
  scale_fill_manual(values = c( "#CEAB07", #NE
                               "#F3DF6C", #MN
                               "#FBA72A", #MO
                               "#FAD510", #IL
                               "#85D4E3",  #VA
                               "#7294D4", #DE
                               "#FAD510", #IL
                               "#85D4E3", #VA
                               "#85D4E3", #VA
                               "#046C9A", #MD
                               "#CEAB07", 
                               "#CEAB07", #NE
                               "#FBA72A", #MO
                               "#85D4E3", #VA
                               "#C52E19", #LA
                               "#7294D4",#DE
                               "#C52E19", #LA
                               "black" #population
                               ))+
  
  labs(x = "% reduction in weed density",
       y = "Site-Year") +
  
  theme_light() +
  theme(
       axis.text.x = element_text(size=8), 
        axis.text.y = element_text(size=7.5, margin = margin(r = -0.5)),
        axis.title.y = element_text(size=10, margin = margin(r = -0.5)), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))
figure2b1
ggsave("percentdiff2mgall.png", width = 2.5, height = 2.5, units = "in", dpi = 300)
```


```{r}

#final_data$ccbm_sc <- scale(final_data$ccbm_mg)
#final_data$DPP_sc  <- scale(final_data$DPP)
#the model
#mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)

ccbm_mean <- mean(final_data$ccbm_mg, na.rm = TRUE)
ccbm_sd   <- sd(final_data$ccbm_mg, na.rm = TRUE)

DPP_mean <- mean(final_data$DPP, na.rm = TRUE)
DPP_sd   <- sd(final_data$DPP, na.rm = TRUE)

ccbm_vals <- c(2) #common highest mean among sites yikes
ccbm_sc_vals <- (ccbm_vals - ccbm_mean) / ccbm_sd #un-scale 

DPP_val <- -14 #focus on 2 weeks pre planting
DPP_sc_val <- (DPP_val - DPP_mean) / DPP_sd #un-scale

model_sy <- unique(mod2@frame$sy) #site year

newdata <- expand.grid(
  sy = model_sy,
  ccbm_sc = ccbm_sc_vals,
  DPP_sc = DPP_sc_val
)

newdata$biomass <- rep(c("2 Mg/ha"), #yikes
                       each = length(model_sy))

#check we didn't remove anything by accident
#length(unique(mod2@frame$sy)) #17 sy
#length(unique(newdata$sy))# 17 sy

newdata <- subset(newdata, !(newdata$sy %in% c("LA.2020", "LA.2019", "MN.2019", "NE.2020", "NE.2019","IL.2019", "DE.2018","MO.2020", "MO.2019", "VA.2019")))

newdata$fit_sy <- predict(mod2, newdata = newdata, re.form = NULL)
newdata$fit_pop <- predict(mod2, newdata = newdata, re.form = NA)
pred_pop <- predictInterval(
  mod2,
  newdata = newdata,
  level = 0.95,
  n.sims = 2000,
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"
)
# convert to %
newdata$percent_sy  <- (1 - exp(newdata$fit_sy)) * 100
newdata$percent_pop <- (1 - exp(newdata$fit_pop)) * 100

newdata$pop_lwr <- (1 - exp(pred_pop$upr)) * 100
newdata$pop_upr <- (1 - exp(pred_pop$lwr)) * 100

pop_data <- data.frame(
  sy = "Population",
  percent_pop = unique(newdata$percent_pop)[1],
  pop_lwr = unique(newdata$pop_lwr)[1],
  pop_upr = unique(newdata$pop_upr)[1]
)

pop_mean <- unique(newdata$percent_pop)[1]

plot_data <- newdata %>%
  mutate(
    deviation = percent_sy - pop_mean
  )

plot_data <- plot_data %>%
  arrange(percent_sy)

plot_data$sy <- factor(plot_data$sy, levels = plot_data$sy)
```


```{r}


figure2b1A <- ggplot(plot_data, aes(y = sy, fill = sy, color = sy)) +
  
  # segment from population mean to site-year
  geom_segment(aes(x = pop_mean, xend = percent_sy,
                   yend = sy),
               color = "gray20", size = 0.5) +
  
  # site-year points
  geom_point(aes(x = percent_sy),
             shape = 21, size = 1.75, stroke = 0.8) +
  
  # population reference line
  geom_vline(xintercept = pop_mean,
             linetype = "longdash",
             color = "black", 
             size = 0.6) +
  # population reference point
 geom_point(data = pop_data,
           aes(x = percent_pop, y = sy),
           shape = 21,
           size = 3,
           fill = "black",
           color = "black") + 
  # population reference point error bar
geom_errorbarh(data = pop_data,
               aes(xmin = pop_lwr,
                   xmax = pop_upr,
                   y = sy),
               height = 0.5,
               color = "black",
               size = 0.5) +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "black",
             size = 0.25) +
    
  scale_color_manual(values = c(
                                "#A18902", #IL
                                "#266B7E", #VA
                                "#266B7E",#VA
                                "#023146",#MD
                                "#9D8206",#NE
                                "#266B7E",#VA
                                "#023146",#DE
                                
                                "black"))+
  scale_fill_manual(values = c( 
                               "#FAD510", #IL
                               "#85D4E3", #VA
                               "#85D4E3", #VA
                               "#046C9A", #MD
                               "#CEAB07", #NE
                               "#85D4E3", #VA
                               "#7294D4",#DE
                              
                               "black" #population
                               ))+
  
  labs(x = "% reduction in weed density",
       y = "Site-Year") +
  
  theme_light() +
  theme(
       axis.text.x = element_text(size=8), 
        axis.text.y = element_text(size=7.5, margin = margin(r = -0.5)),
        axis.title.y = element_text(size=10, margin = margin(r = -0.5)), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))
figure2b1A
ggsave("percentdiff2mgsubset.png", width = 5, height = 2.5, units = "in", dpi = 300)
```




```{r}

## model
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)


##what I had
eff1 <- effect("DPP_sc", mod2, partial.residuals = T)
closest <- function(x, x0) apply(outer(x, x0, FUN = function(x, x0) abs(x - x0)), 1, which.min)
x.fit <- unlist(eff1$x.all)
trans <- I

x <- data.frame(lower = eff1$lower, upper = eff1$upper, fit = eff1$fit, DPP_sc = eff1$x$DPP_sc)
xy <- data.frame(x = x.fit, y = x$fit[closest(trans(x.fit), x$DPP_sc)] + eff1$residuals)


lnrr_dpp_sc_effects <- ggplot(x, aes(x = DPP_sc, y = fit)) +
  geom_point(data = xy, aes(x = x, y = y),
            shape = 16, col = "gray", size = 2, alpha = 0.25)+

 geom_line(data = x, aes(x = DPP_sc, y = fit),
          size = 1.2, col = "gray40")+
 # geom_point(data = xy, aes(x = x, y = y), shape = 16, col = "#7ecdb5", size = 2, alpha = 0.25) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.15) +
#  geom_smooth(data = xy, aes(x = trans(x), y = y), 
 #             method = "lm", span = 2/3, linetype = "solid", se = FALSE, col = "#E6A0C4", size = 1.2) +
  geom_hline(yintercept = 0, lty = "dashed", size = 1)+
labs(
  x = "Cover crop termination timing",
  y = "Weed density (4 WAP)") + # not convinced I like this label
 theme_light() +
  theme(
       axis.text.x = element_text(size=8), 
        axis.text.y = element_text(size=7.5, margin = margin(r = -0.5)),
        axis.title.y = element_text(size=10, margin = margin(r = -0.5)), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))
lnrr_dpp_sc_effects
ggsave("ttondensity.png", width = 2.5, height = 2.5, units = "in", dpi = 300)
```



```{r}
dat <- na.omit(final_data[, c("lnRR_4WAP", "ccbm_sc", "DPP_sc")])

y_resid <- resid(lm(lnRR_4WAP ~ ccbm_sc, data = dat))
x_resid <- resid(lm(DPP_sc ~ ccbm_sc, data = dat))

plot(x_resid, y_resid)

plot <- ggplot(dat, aes(x = x_resid, y = y_resid)) +
  geom_point(data = dat, aes(x = x_resid, y = y_resid),
            shape = 16, col = "gray", size = 2, alpha = 0.25)+


#geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.15) +
  geom_hline(yintercept = 0, lty = "dashed", size = 0.5)+
  geom_smooth(method = "lm", se = TRUE, 
          size = 1.2, col = "gray45")+
labs(
  x = "Cover crop termination timing",
  y = "Weed density (LRR; 4 WAP)") + # not convinced I like this label
 theme_light() +
  theme(
       axis.text.x = element_text(size=8), 
        axis.text.y = element_text(size=7.5, margin = margin(r = -0.5)),
        axis.title.y = element_text(size=10, margin = margin(r = -0.5)), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))
plot
ggsave("residuals.png", width = 2.5, height = 2.5, units = "in", dpi = 300)
```


## Temporal dynamics

I will be repeating the above process for the next pSEM which includes mid- and end- of season weed density

There will be 3 component models

3. cover crop biomass ~ termination timing
2. Weed density mid season ~ cover crop biomass
1. Weed density end season ~ Weed density mid season

Removing specific outliers
```{r}
#x11()
#plot(final_data$DPP, final_data$lnRR_harvest, pch=19, cex=2, col="goldenrod")

# identify points
#idx <- identify(final_data$DPP, final_data$lnRR_harvest, labels = final_data$plotID)

# confirm
#final_data[idx, c("plotID", "DPP", "lnRR_harvest")]

# remove by row

final_data <- subset(final_data, !(final_data$plotID %in% c("L92LW", "D93LW", "M86LW", "M04LW")))

ggplot(final_data, aes(x = DPP, y = lnRR_4WAP, color = state)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_smooth(method = "lm", fullrange = TRUE, size = 1, alpha = 0.2) +
  theme_light()
```


### Model 3 (cover crop biomass ~ days pre planting)

Same as above model 1

```{r eval = FALSE}
mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod1a <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc || sy), data = final_data) # uncorrelated random effects
anova(mod1a, mod1) # correlated is better!
summary(mod1)
```

### Model 4 (weeds mid ~ cover crop biomass)

Same as above model 2


```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit) #slope correlated
mod2a <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data) # uncorrelated random effects
anova(mod2a, mod2) # correlated is better!

mod2b <- lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data) # remove dpp
anova(mod2b, mod2) # on the edge here

summary(mod2)
```

### Model 5 (weeds end ~ weeds mid)


```{r}
mod5 <- lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit)
#slope correlated
#singularity
anova(mod5) # non significant ccbm fixed
mod5b <- lmer(lnRR_harvest ~ lnRR_4WAP + DPP_sc + (ccbm_sc | sy),data = final_data) # remove ccbm
anova(mod5b, mod5) # no difference
anova(mod5b)

mod5 <- lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy),data = final_data) 
mod5a <- lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data) # uncorrelated random effects
#no singularity
anova(mod5a, mod5) # correlated is better!
#except we're still having singularities (but not lack of convergence)

summary(mod5)
```


```{r   fig.width = 5, fig.height = 5}
residuals <- resid(mod5)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod5, type="link"), pearson = residuals(mod5, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod5, vars = "Cook")
dotplot(ranef(mod5))
```


```{r, fig.height = 15, fig.width = 10}
re <- ranef(mod5a)$sy
re_df <- re %>%
  tibble::rownames_to_column("sy") %>%
  rename(rand_intercept = `(Intercept)`)
fix_int <- fixef(mod5a)[1]
re_df <- re_df %>%
  mutate(intercept = fix_int + rand_intercept)
final_data1 <- final_data %>%
  left_join(re_df, by = "sy")
final_data1 <- final_data1 %>%
  mutate(sy = reorder(sy, intercept))

ggplot(final_data1, aes(x = ccbm_sc.x, y = lnRR_harvest)) +
  geom_point(size = 2, alpha = 0.2, pch = 16) +
  geom_smooth(method = "lm", se = TRUE,
              fullrange = TRUE, 
              size = 0.75, alpha = 0.5,
              col = "seagreen1") +
  theme_light() + 
  theme(axis.title = element_text(size = 14),
        legend.title = element_text(size = 14), 
        aspect.ratio = 1.5)+
  labs(y = "lnRR_4WAP", x = "ccbmscaled")+
  facet_wrap(~sy)
```


```{r, fig.height = 15, fig.width = 10}
re <- ranef(mod5a)$sy
re_df <- re %>%
  tibble::rownames_to_column("sy") %>%
  rename(rand_intercept = `(Intercept)`)
fix_int <- fixef(mod5a)[1]
re_df <- re_df %>%
  mutate(intercept = fix_int + rand_intercept)
final_data1 <- final_data %>%
  left_join(re_df, by = "sy")
final_data1 <- final_data1 %>%
  mutate(sy = reorder(sy, intercept))

ggplot(final_data1, aes(x = DPP_sc, y = lnRR_harvest)) +
  geom_point(size = 2, alpha = 0.2, pch = 16) +
  geom_smooth(method = "lm", se = TRUE,
              fullrange = TRUE, 
              size = 0.75, alpha = 0.5,
              col = "seagreen1") +
  theme_light() + 
  theme(axis.title = element_text(size = 14),
        legend.title = element_text(size = 14), 
        aspect.ratio = 1.5)+
  labs(y = "lnRR_4WAP", x = "ccbmscaled")+
  facet_wrap(~sy)
```


### PSEM

Starting-- most basic, no end-season density

DPP -> ccbm 
ccbm -> lnrr_mid
lnrr_mid -> lnrr_end
```{r}
psemDensity2 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity2, .progressBar = FALSE)
#  p values v small
fisherC(psemDensity2)
#Fisher's: 18.514	 on 6 df, p = 0.005
summary(psemDensity2)
```


DPP -> ccbm 
ccbm + DPP -> lnrr_mid
lnrr_mid -> lnrr_end
```{r}
psemDensity3 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity3, .progressBar = FALSE)
# still small p value
fisherC(psemDensity3)
# Fisher's C 9.271	on 4 df p = 0.055
summary(psemDensity3)
```

DPP -> ccbm 
ccbm -> lnrr_mid
lnrr_mid + ccbm -> lnrr_end
```{r}
psemDensity4 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity4, .progressBar = FALSE)
# still small p value
fisherC(psemDensity4)
# Fisher's C 11.647	on 4 df p = 0.02	
summary(psemDensity4)
```


DPP -> ccbm 
ccbm + DPP -> lnrr_mid
lnrr_mid + ccbm -> lnrr_end
```{r}
psemDensity5 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity5, .progressBar = FALSE)
# still small p value
fisherC(psemDensity5)
# Fisher's C 6.286	on 2 df p = 0.043		
summary(psemDensity5)
```

DPP -> ccbm 
ccbm + DPP -> lnrr_mid
lnrr_mid + ccbm + DPP -> lnrr_end
```{r}
psemDensity6 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity6, .progressBar = FALSE)
# saturated
fisherC(psemDensity6)
# saturated
summary(psemDensity6)
```


DPP -> ccbm 
ccbm + DPP -> lnrr_mid
lnrr_mid + ccbm + DPP -> lnrr_end
```{r}
psemDensity7 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc +(ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity7, .progressBar = FALSE)
# saturated
fisherC(psemDensity7)
# saturated
summary(psemDensity7)
```


Ok here both of the lnrr models have small p values but the harvest one is smaller p = 0.03 vs p = 0.05... so I'm coming back to here and will add dpp to the top model (harvest)
```{r}
psemDensity4 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity4, .progressBar = FALSE)
# still small p value
fisherC(psemDensity4)
# Fisher's C 12.729	on 4 df p = 0.013	
summary(psemDensity4)
```


```{r}
psemDensity8 <- psem(
  lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_data, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_data, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_data, na.action = na.omit)
)

dSep(psemDensity8, .progressBar = FALSE)
# 0.05
fisherC(psemDensity8)
# Fisher's C 5.924	on 2 df p = 0.05	
summary(psemDensity8)
```







## Cash crop presence


Soybean yield

Starting everything from scratch

```{r warning = FALSE, include = FALSE}
#setwd("~/CCHIT 2025")
cchit <- read_csv("biomass_master.csv") 
cchit <- as.data.frame(cchit)

cchit$state <- as.factor(cchit$state)
cchit$year <- as.factor(cchit$year)
cchit$plot <- as.factor(cchit$plot)
cchit$term.trmt <- as.factor(cchit$term.trmt)
cchit$herb.trmt <- as.factor(cchit$herb.trmt)
cchit$region <- as.factor(cchit$region)
cchit$DPP <- as.numeric(cchit$DPP) 
cchit$sy <- interaction(cchit$state, cchit$year) 
cchit$plotID <- as.factor(cchit$plotID)

str(cchit)

cchit <- cchit %>% dplyr::rename(ccbm = "cc.bm.kgha", yield = "yield.kgha", weedbmh = "weed.bm.harv.kgha", weedbmt = "weed.bm.term.kgha", densityh = "density.harv.allspp.m2", densityt = "density.term.allspp.m2", sbpop = "soybean.pop.ha") 

cchit$weedbmh_mg <- as.numeric(cchit$weedbmh/1000) 
cchit$weedbmt_mg <- as.numeric(cchit$weedbmt/1000)
cchit$ccbm_mg <- as.numeric(cchit$ccbm/1000)

cchith <- subset(cchit, herb.trmt == "Weedy") # subset 
cchith$yield_mg <- as.numeric(cchith$yield/1000)
cchith <- cchith %>%
 dplyr:: mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

group_means <- cchith %>%
dplyr::  group_by(state, year, sy, term.trmt, plotID, DPP, ccbm_mg) %>%
  dplyr::summarise(mean_yield = mean(yield_mg, na.rm = TRUE), .groups = "drop")

group_means <- group_means %>%
  dplyr::mutate(block = substr(plotID, 2, 3))

ref <- group_means %>%
  dplyr::filter(term.trmt == "No cover") %>%
  dplyr::rename(ref_biomass = mean_yield) %>%
  dplyr::select(sy, block, ref_biomass)

rr_yield <- group_means %>%
  left_join(ref, by = c("sy", "block")) %>%
  mutate(
    RR = mean_yield / ref_biomass,
    lnyield = log(mean_yield + 1e-6)
  ) %>%
  filter(!is.na(RR), is.finite(RR))

  
rr_yield[sapply(rr_yield, is.infinite)] <- NA
rr_yield <- na.omit(rr_yield)


rr_yield$RR <- as.numeric(rr_yield$RR)
rr_yield <- as.data.frame(rr_yield)

rr_yield$lnyield <- log(rr_yield$mean_yield + 0.000001) 

str(rr_yield)

```


Density at harvest

```{r warning = FALSE}
cchitnh <- subset(cchit, herb.trmt == "Weedy") # subset out herbicide treated
cchitnh <- cchitnh %>%
  dplyr::mutate(across(c(sy, plotID, term.trmt, herb.trmt), as.factor))

group_means <- cchitnh %>%
  dplyr::group_by(state, year, sy, term.trmt, plotID, DPP, ccbm_mg) %>%
  dplyr::summarize(
    mean_biomass = mean(densityh, na.rm = TRUE),
    .groups = "drop"
  ) 

ref <- group_means %>%
  dplyr::filter(term.trmt == "No cover") %>%
  dplyr::rename(ref_biomass = mean_biomass) %>%
   dplyr :: select(sy, plotID, ref_biomass)

group_means <- group_means %>%
  dplyr::mutate(block = substr(plotID, 2, 3))  
ref <- ref %>%
  dplyr::mutate(block = substr(plotID, 2, 3))

rr_data <- group_means %>%
  dplyr::left_join(ref, by = "sy", suffix = c("", "_ref")) %>%
  dplyr::filter(!is.na(ref_biomass), !is.na(mean_biomass)) %>%
  dplyr::mutate(RR = mean_biomass / ref_biomass) %>%
  dplyr::select(-plotID_ref, -block_ref) 

rr_data[sapply(rr_data, is.infinite)] <- NA
rr_data <- na.omit(rr_data)


rr_data$RR <- as.numeric(rr_data$RR)
rr_data <- as.data.frame(rr_data)
str(rr_data)

rr_data$lnrr <- log(rr_data$RR + 0.000001) 
```



Making a new dataset with all combined

```{r}
#rr_yield <- rr_yield %>%
 #rename(lnyield = lnrr)

#rr_density <- rr_density %>%
 # rename(lnRR_4WAP = lnRR)

rr_data <- rr_data %>%
  dplyr::rename(lnRR_harvest = lnrr)

rr_yield <- rr_yield %>%
  dplyr::mutate(across(c(sy, plotID, block), as.character))

#rr_density <- rr_density %>%
 # mutate(across(c(sy, plotID, block), as.character))

rr_data <- rr_data %>%
  dplyr::mutate(across(c(sy, plotID, block), as.character))

rr_density_mean <- rr_density %>%
  dplyr::group_by(sy, block, DPP, ccbm_mg) %>%
  dplyr::summarise(lnRR_4WAP = mean(lnRR_4WAP, na.rm = TRUE), .groups = "drop")

rr_yield_mean <- rr_yield %>%
  dplyr::group_by(sy, block, DPP, ccbm_mg) %>%
  dplyr::summarise(lnyield = mean(lnyield, na.rm = TRUE), .groups = "drop")

final_datay <- rr_data %>%
  left_join(rr_density_mean, by = c("sy", "block", "DPP", "ccbm_mg"))
final_datay <- final_datay %>%
  left_join(rr_yield_mean, by = c("sy", "block", "DPP", "ccbm_mg"))

str(final_datay)
```




END OF SEASON WEEDS

### Model 9
```{r}
final_datay$ccbm_sc <- as.numeric(scale(final_datay$ccbm_mg))
final_datay$DPP_sc  <- as.numeric(scale(final_datay$DPP))
```

```{r}
mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
```

## Model 10
```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit)
```

## Model 11 (lnrr harvest ~ lnrr mid)

```{r}
mod5a <- lmer(lnRR_harvest ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)
```

## Model 11.5 (lnrr harvest ~ ccbm)

Per previous suggestions not doing both weed density timings 

```{r}
mod11_5 <- lmer(lnRR_harvest ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)
mod11_5a <- lmer(lnRR_harvest ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
anova(mod11_5, mod11_5a) # correlated is way better!
summary(mod11_5a)
anova(mod11_5a)
```

## Model 11.75 (lnRR_4WAP ~ ccbm)

Per previous suggestions not doing both weed density timings 

```{r}
mod11_75 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)
mod11_75a <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
anova(mod11_75, mod11_75a) # correlated is way better!
summary(mod11_75a)
anova(mod11_75a)
```




## Model 12.5 (soybean yield ~ lnrr end)

```{r}
mod9a <- lmer(lnyield ~ lnRR_harvest + lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
anova(mod9a) #biggest p value is 4wap
mod9b <- lmer(lnyield ~ lnRR_harvest +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
mod9c <- lmer(lnyield ~ lnRR_harvest +  ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)

anova(mod9b, mod9c)
```

P-value >0.05 as such, best model is the most parsimonious.


```{r   fig.width = 5, fig.height = 5}
residuals <- resid(mod9c)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod9c, type="link"), pearson = residuals(mod9c, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod9c, vars = "Cook")
dotplot(ranef(mod9c))
```





## Model 12 (soybean yield ~ lnrr 4wap)

```{r}
mod9a <- lmer(lnyield ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
anova(mod9a) #biggest p value is 4wap
mod9b <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
mod9c <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)

anova(mod9b, mod9c)
```

P-value >0.05 as such, best model is the most parsimonious.




```{r}
Q1 <- quantile(residuals, 0.25)
Q3 <- quantile(residuals, 0.75)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
outliers <- which(residuals < lower_bound | residuals > upper_bound)

final_datay <- final_datay[-outliers, ]
```

```{r}
mod9c <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)
residuals <- resid(mod9c)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod9c, type="link"), pearson = residuals(mod9c, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod9c, vars = "Cook")
dotplot(ranef(mod9c))
```




### PSEM

Starting-- most basic, no mid-season density


```{r}
psemDensity2 <- psem(
  lmer(lnyield ~ lnRR_harvest + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_harvest ~ ccbm_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity2, .progressBar = FALSE)
# 3 p values v small
fisherC(psemDensity2)
#Fisher's: 94.036	on 6 df, p = 0
summary(psemDensity2)
```



```{r}
psemDensity3 <- psem(
  lmer(lnyield ~ lnRR_harvest + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_harvest ~ ccbm_sc+ (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity3, .progressBar = FALSE)
# still small p value
fisherC(psemDensity3)

summary(psemDensity3)
```


```{r}
psemDensity4 <- psem(
  lmer(lnyield ~ lnRR_harvest + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_harvest ~ ccbm_sc  + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity4, .progressBar = FALSE)
# still small p value
fisherC(psemDensity4)

summary(psemDensity4)
```


```{r}
psemDensity5 <- psem(
  lmer(lnyield ~ lnRR_harvest + ccbm_sc +  DPP_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_harvest ~ ccbm_sc  + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity5, .progressBar = FALSE)
# still small p value
fisherC(psemDensity5)

summary(psemDensity5)
```


```{r}
psemDensity6 <- psem(
  lmer(lnyield ~ lnRR_harvest + ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_harvest ~ ccbm_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity6, .progressBar = FALSE)
# saturated
fisherC(psemDensity6)
# saturated
summary(psemDensity6)
```


```{r}
psemDensity6 <- psem(
  lmer(lnyield ~ lnRR_harvest + ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_harvest ~ ccbm_sc + DPP_sc+ (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity6, .progressBar = FALSE)
# saturated
fisherC(psemDensity6)
# saturated
summary(psemDensity6)
```



## PSEM 
Mid season this time ig


```{r}
psemDensity2 <- psem(
  lmer(lnyield ~ lnRR_4WAP + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity2, .progressBar = FALSE)
# 3 p values v small
fisherC(psemDensity2)
#Fisher's: 74.059	on 6 df, p = 0
summary(psemDensity2)
```



```{r}
psemDensity3 <- psem(
  lmer(lnyield ~ lnRR_4WAP + ccbm_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc+ (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity3, .progressBar = FALSE)
# still small p value
fisherC(psemDensity3)
# 79.501	 4df p < 0.001
summary(psemDensity3)
```


```{r}
psemDensity4 <- psem(
  lmer(lnyield ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity4, .progressBar = FALSE)
# still small p value
fisherC(psemDensity4)
# 7.314 2df p = 0.017
summary(psemDensity4)

basisSet(psemDensity4)
dSep(psemDensity4)
```


```{r}
psemDensity5 <- psem(
  lmer(lnyield ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity5, .progressBar = FALSE)

fisherC(psemDensity5)

summary(psemDensity5)
```


### Figures

```{r}
mod9b <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)


valid_sy <- unique(mod9c@frame$sy) #just site years in mod2

newdat <- expand.grid(
  ccbm_sc = seq(min(final_datay$ccbm_sc, na.rm = TRUE),
                max(final_datay$ccbm_sc, na.rm = TRUE),
                length.out = 100),
  DPP_sc = mean(final_datay$DPP_sc, na.rm = TRUE),
  lnRR_4WAP = mean(final_datay$lnRR_4WAP, na.rm = TRUE), 
  sy = valid_sy
)
newdat$pred_sy <- predict(mod9b, newdata = newdat, re.form = NULL)

newdat$pred_pop <- predict(mod9b, newdata = newdat, re.form = NA)

pred_int2 <- predictInterval(
  mod9b,
  newdata = newdat,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat$fit   <- pred_int2$fit
newdat$lwr   <- pred_int2$lwr
newdat$upr   <- pred_int2$upr

figure5b <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_sy, color = sy),
            size = 0.75, alpha = 0.4) +
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat,
            aes(x = ccbm_sc, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
  labs(x = "Cover crop biomass",
       y = "Soybean Yield", 
       color = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#7294D4", "#7294D4",  #delaware
                                "#FAD510", "#FAD510", #illinois
                                "#C52E19",  #louisiana
                                "#046C9A", #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A", #missouri
                                "#CEAB07", "#CEAB07", "#CEAB07",#nebraska
                                
                                "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(5, "pt"),
       legend.text = element_text(size = 7),
       legend.title = element_text(size = 8),
       legend.key.size = unit(0.2, "cm"),
       legend.key.spacing.y = unit(0.1, "cm"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE, title.position = "top", title.hjust = 0.5))

figure5b

#work out some save details 
ggsave("figure5biomass.png", width = 2.5, height = 2, units = "in", dpi = 300)
```



```{r}
mod9b <- lmer(lnyield ~ lnRR_harvest + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)


valid_sy <- unique(mod9b@frame$sy) #just site years in mod2

newdat <- expand.grid(
  ccbm_sc = seq(min(final_datay$ccbm_sc, na.rm = TRUE),
                max(final_datay$ccbm_sc, na.rm = TRUE),
                length.out = 100),
  DPP_sc = mean(final_datay$DPP_sc, na.rm = TRUE),
  lnRR_harvest = mean(final_datay$lnRR_harvest, na.rm = TRUE), 
  #lnRR_4WAP = mean(final_datay$lnRR_4WAP, na.rm = TRUE),
  sy = valid_sy
)
newdat$pred_sy <- predict(mod9b, newdata = newdat, re.form = NULL)

newdat$pred_pop <- predict(mod9b, newdata = newdat, re.form = NA)

pred_int2 <- predictInterval(
  mod9b,
  newdata = newdat,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat$fit   <- pred_int2$fit
newdat$lwr   <- pred_int2$lwr
newdat$upr   <- pred_int2$upr

figure5b_end <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_sy, color = sy),
            size = 0.75, alpha = 0.4) +
  geom_line(data = newdat,
            aes(x = ccbm_sc, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat,
            aes(x = ccbm_sc, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
  labs(x = "Cover crop biomass",
       y = "Soybean Yield", 
       color = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#7294D4", "#7294D4",   #delaware
                                "#FAD510", "#FAD510", #illinois
                                "#C52E19",  #louisiana
                                "#046C9A",  #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A", #missouri
                                "#CEAB07", "#CEAB07", "#CEAB07", #nebraska
                                
                                "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(5, "pt"),
       legend.text = element_text(size = 7),
       legend.title = element_text(size = 8),
       legend.key.size = unit(0.2, "cm"),
       legend.key.spacing.y = unit(0.1, "cm"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE, title.position = "top", title.hjust = 0.5))

figure5b_end

#work out some save details 
ggsave("figure_endseason.png", width = 2.5, height = 2, units = "in", dpi = 300)
```










MID SEASON WEEDS

### Model 9
```{r}
final_datay$ccbm_sc <- as.numeric(scale(final_datay$ccbm_mg))
final_datay$DPP_sc  <- as.numeric(scale(final_datay$DPP))
```

```{r}
mod1 <- lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
```

## Model 10
```{r}
mod2 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit)
```

## Model 11 (lnrr harvest ~ lnrr mid)

```{r}
mod5a <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)
```

# Model 11.5 (lnrr harvest ~ ccbm)

Per previous suggestions not doing both weed density timings 

```{r}
mod11_5 <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)
mod11_5a <- lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
anova(mod11_5, mod11_5a) # correlated is way better!
summary(mod11_5a)
anova(mod11_5a)
```

## Model 12 (soybean yield ~ lnrr end)

```{r}
mod9a <- lmer(lnyield ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
anova(mod9a) 
mod9b <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)
mod9c <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)

anova(mod9b, mod9c)
```

P-value >0.05 as such, best model is the most parsimonious.



```{r   fig.width = 5, fig.height = 5}
residuals <- resid(mod9c)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod9c, type="link"), pearson = residuals(mod9c, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod9c, vars = "Cook")
dotplot(ranef(mod9c))
```

```{r}
Q1 <- quantile(residuals, 0.25)
Q3 <- quantile(residuals, 0.75)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
outliers <- which(residuals < lower_bound | residuals > upper_bound)

final_datay <- final_datay[-outliers, ]
```

```{r}
mod9c <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay)
residuals <- resid(mod9c)
hist(residuals, breaks = 20, main = "Histogram of Residuals")
ggplot(data.frame(eta = predict(mod9c, type="link"), pearson = residuals(mod9c, type = "pearson")),
    aes(x = eta,y = pearson)) +
    geom_point() +
    theme_bw()
qqnorm(residuals)
qqline(residuals, col="red")
infIndexPlot(mod9c, vars = "Cook")
dotplot(ranef(mod9c))
```




### PSEM

Starting-- most basic, no mid-season density


```{r}
psemDensity2 <- psem(
  lmer(lnyield ~ lnRR_4WAP + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity2, .progressBar = FALSE)
# 3 p values v small
fisherC(psemDensity2)

summary(psemDensity2)
```



```{r}
psemDensity3 <- psem(
  lmer(lnyield ~ lnRR_4WAP + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity3, .progressBar = FALSE)
# still small p value
fisherC(psemDensity3)

summary(psemDensity3)
```


```{r}
psemDensity4 <- psem(
  lmer(lnyield ~ lnRR_4WAP + ccbm_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity4, .progressBar = FALSE)
# still small p value
fisherC(psemDensity4)

summary(psemDensity4)
```


```{r}
psemDensity5 <- psem(
  lmer(lnyield ~ lnRR_4WAP + ccbm_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity5, .progressBar = FALSE)
# still small p value
fisherC(psemDensity5)

summary(psemDensity5)
```


```{r}
psemDensity6 <- psem(
  lmer(lnyield ~ lnRR_4WAP + ccbm_sc + DPP_sc + (ccbm_sc || sy), data = final_datay, na.action = na.omit),
  lmer(lnRR_4WAP ~ ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay, na.action = na.omit),
  lmer(ccbm_sc ~ DPP_sc + (DPP_sc | sy), data = final_datay, na.action = na.omit)
)

dSep(psemDensity6, .progressBar = FALSE)
# saturated
fisherC(psemDensity6)
# saturated
summary(psemDensity6)
```


### Figures





```{r}
mod9b <- lmer(lnyield ~ lnRR_4WAP +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)


valid_sy <- unique(mod9c@frame$sy) #just site years in mod2

newdat <- expand.grid(
  DPP_sc = seq(min(final_datay$DPP_sc, na.rm = TRUE),
                max(final_datay$DPP_sc, na.rm = TRUE),
                length.out = 100),
  ccbm_sc = mean(final_datay$ccbm_sc, na.rm = TRUE),
  lnRR_4WAP = mean(final_datay$lnRR_4WAP, na.rm = TRUE), 
  sy = valid_sy
)
newdat$pred_sy <- predict(mod9b, newdata = newdat, re.form = NULL)

newdat$pred_pop <- predict(mod9b, newdata = newdat, re.form = NA)

pred_int2 <- predictInterval(
  mod9b,
  newdata = newdat,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat$fit   <- pred_int2$fit
newdat$lwr   <- pred_int2$lwr
newdat$upr   <- pred_int2$upr

figure5c <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat,
            aes(x = DPP_sc, y = pred_sy, color = sy),
            size = 0.75, alpha = 0.4) +
  geom_line(data = newdat,
            aes(x = DPP_sc, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat,
            aes(x = DPP_sc, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
  labs(x = "Termination timing",
       y = "Soybean Yield", 
       color = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#7294D4", "#7294D4",  #delaware
                                "#FAD510", "#FAD510", #illinois
                                "#C52E19",  #louisiana
                                "#046C9A", #maryland
                                "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A", #missouri
                                "#CEAB07", "#CEAB07", "#CEAB07",#nebraska
                                
                                "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
       legend.position = "none",
       legend.spacing.x = unit(10, "pt"),
       legend.text = element_text(size = 8),
       legend.title = element_text(size = 9),
       legend.key.size = unit(0.5, "cm"),
       legend.key.spacing.y = unit(0, "pt"),
       legend.box.margin = margin(-5, -5, -5, -5))+
  guides(col = guide_legend(nrow = 3, byrow = TRUE))


figure5c

ggsave("figure5termtiming.png", width = 3, height = 2.5, units = "in", dpi = 300)
```


```{r}
mod9b <- lmer(lnyield ~ lnRR_harvest +  ccbm_sc + DPP_sc + (ccbm_sc | sy), data = final_datay)


valid_sy <- unique(mod9c@frame$sy) #just site years in mod2

newdat <- expand.grid(
  lnRR_harvest = seq(min(final_datay$lnRR_harvest, na.rm = TRUE),
                max(final_datay$lnRR_harvest, na.rm = TRUE),
                length.out = 100),
  DPP_sc = mean(final_datay$DPP_sc, na.rm = TRUE),
  ccbm_sc = mean(final_datay$ccbm_sc, na.rm = TRUE), 
  sy = valid_sy
)
newdat$pred_sy <- predict(mod9b, newdata = newdat, re.form = NULL)

newdat$pred_pop <- predict(mod9b, newdata = newdat, re.form = NA)

pred_int2 <- predictInterval(
  mod9b,
  newdata = newdat,
  level = 0.95, #seems fine
  n.sims = 1000, #runs fast?? good enough
  stat = "mean",
  include.resid.var = FALSE,
  which = "fixed"  # we only want the confidence interval for the population line not the site year lines
)

newdat$fit   <- pred_int2$fit
newdat$lwr   <- pred_int2$lwr
newdat$upr   <- pred_int2$upr

figure5d <- ggplot() +
  # Site-year lines (colored)
  geom_line(data = newdat,
            aes(x = lnRR_harvest, y = pred_sy, color = sy),
            size = 0.75, alpha = 0.4) +
  geom_line(data = newdat,
            aes(x = lnRR_harvest, y = pred_pop),
            color = "black", size = 1.5) +
  geom_ribbon(data = newdat,
            aes(x = lnRR_harvest, ymin = lwr, ymax = upr),
            fill = "gray30", alpha = 0.4) +
  labs(x = "lnRR_harvest",
       y = "Soybean Yield (natural logarithmically transformed)", 
       color = "Site-year") + 
  theme_light () + 
  scale_color_manual(values = c("#7294D4", "#7294D4", "#7294D4", "#7294D4", #delaware
                                "#FAD510", "#FAD510", "#FAD510", #illinois
                                "#C52E19",  #louisiana
                                "#046C9A", "#046C9A", "#046C9A", #maryland
                                "#F3DF6C", "#F3DF6C", "#F3DF6C", "#F3DF6C", #minnesota
                                "#FBA72A", "#FBA72A", "#FBA72A", #missouri
                                "#CEAB07", "#CEAB07", "#CEAB07", "#CEAB07",#nebraska
                                "#972D15", #texas
                                "#85D4E3", "#85D4E3", "#85D4E3")) + #virginia
 
   
  theme(
       axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=10), 
        axis.title.x = element_text(size=10),
        legend.position = "none")

figure5d

#work out some save details 
#ggsave("figure1bottom.png", width = 6, height = 2.5, units = "in", dpi = 300)
```





